---
title: "HBSC: Modeling"
date: "Updated: `r format(Sys.time(), '%Y-%m-%d')`"
output:
  pdf_document: default
  word_document: default
  html_notebook: default
urlcolor: blue
---

```{r}

rm(list = ls())

setwd("C:/MisLocalFiles/Github/HBSC/scripts")
project_folder <- "C:/MisLocalFiles/Github/HBSC/"

require(tidyverse)
require(caret)
require(broom)

```

# Data Overview
This data is prepared in hbsc.rmd file. It's dimensions are:

```{r}
# ---------------------------------------------------------
# load dataset
# ---------------------------------------------------------

d <- read_csv(paste0(project_folder, "data/processed/dat_2026-08-16.csv"))

dim(d) #244097     25

```

The list of vars and it's type is:

```{r}

glimpse(d)

```

# Missing Values
Most features have missing values. 

```{r}

summary(d)

```

For now, I will just keep complete observations to simplify my life. 
But as can be seen some countries are very afected. Netherlands is a good option.

```{r}

d <- d %>% 
  mutate(any_na = if_any(everything(), is.na))

d %>%
  group_by(country_name, any_na) %>%
  tally() %>%
  pivot_wider(names_from = any_na,
              names_prefix = "na_",
              values_from = n,
              values_fill = 0) %>%
  mutate(n = na_FALSE + na_TRUE) %>%
  mutate(pct_complete = round(na_FALSE / n,3)) %>%
  select(n, na_FALSE, pct_complete) %>%
  arrange(desc(pct_complete))-> tbl_nas_country

kable(tbl_nas_country)

dc <- d[complete.cases(d),]

```

# Check if Dataset if balanced based on vars to predict

## lifesat_low
For lifesat_low I like Canada and Netherlands in the extremes. Could be good to compare
if we opt for two countries. For now, I will go with Canada due to big %.

```{r}

dc %>%
  group_by(country_name, lifesat_low) %>%
  tally() %>%
  pivot_wider(names_from = lifesat_low,
              names_prefix = "is_",
              values_from = n,
              values_fill = 0) %>%
  mutate(n = is_0 + is_1) %>%
  mutate(lifesat_low_pct = round(is_1/n,3)) %>%
  select(country_name, n, lifesat_low_pct) %>%
  arrange(desc(lifesat_low_pct)) -> lifesat_low_pct_country

kable(lifesat_low_pct_country)

#mental_complaintsY   

```

## mental_complaintsY
For mental_complaintsY, Netherlads is still good. But Canada is more halfway,
while other countries like Italy are at top. Israel is at top of list on both
so could try that. 

```{r}

dc %>%
  group_by(country_name, mental_complaintsY) %>%
  tally() %>%
  pivot_wider(names_from = mental_complaintsY,
              names_prefix = "is_",
              values_from = n,
              values_fill = 0) %>%
  mutate(n = is_0 + is_1) %>%
  mutate(mental_complaintsY_pct = round(is_1/n,3)) %>%
  select(country_name, n, mental_complaintsY_pct) %>%
  arrange(desc(mental_complaintsY_pct)) -> mental_complaintsY_pct_country

kable(mental_complaintsY_pct_country)

```

Well, ok. I need to make a call for now. So I will pick Italy first since 
it has the highest % of mental health complaints. And if I decide to compare
we can do it with Spain or Netherlands.  

# Logical grouping of variables
I group logically in themes the 25 vars as follows:

Vars to Predict:
- lifesat_low
- mental_complaintsY (or maybe health_complaintsY)

Key Predictors to Study (in this order):

Problematic social media use: 
- pmsu_lmh (over pmsu_yn)

Other concerning online stuff:
- emconlfreqYes (too much online comms)
- lives_onlineYN (prefer online vulnerability)

Bulling:
- beenbulliedYes
- cbeenbulliedYes

Youth support system:
- family_support_high (over family_support_hml)
- friends_support_high (over friends_support_hml)
- school_support_high (over teacher_support_high:student_support_high)

Parental support:
- talkfatherYes x talkmotherYes

Habits good/bad:
- timeexe_r (exercise)
- alc30d_YN

Covariates:
- age_r
- sex_r
- IRRELFAS_LMH_r

Other:
- seqno_int
- country_name

# Coercions of Data Types
For modeling, many vars need to be coerced so I start there before univ EDA.

```{r}

dc <- dc %>%
  mutate(lifesat_low = factor(lifesat_low, levels = c("0","1")),
         mental_complaintsY = factor(mental_complaintsY, levels = c("0","1")),
         health_complaintsY = factor(health_complaintsY, levels = c("0", "1")),
         pmsu_lmh = factor(pmsu_lmh, levels = c("Low", "Med", "High")),
         pmsu_yn = factor(pmsu_yn, levels = c("0", "1")),
         family_support_high = factor(family_support_high, levels = c("0", "1")),
         family_support_hml = factor(family_support_hml, levels = c("low", "med", "high")),
         friends_support_high = factor(friends_support_high, levels = c("0", "1")),
         friends_support_hml = factor(friends_support_hml, levels = c("low", "med", "high")),
         teacher_support_high = factor(teacher_support_high, levels = c("0", "1")),
         student_support_high = factor(student_support_high, levels = c("0", "1")),
         school_support_high = factor(school_support_high, levels = c("0", "1")),
         IRRELFAS_LMH_r = factor(IRRELFAS_LMH_r, levels = c("Low20", "Med60", "High20")),
         talkfatherYes = factor(talkfatherYes, levels = c("0", "1")),
         talkmotherYes = factor(talkmotherYes, levels = c("0", "1")),
         beenbulliedYes = factor(beenbulliedYes, levels = c("0", "1")),
         cbeenbulliedYes = factor(cbeenbulliedYes, levels = c("0", "1")),
         age_r = factor(age_r, levels = c("11", "13", "15")),
         sex_r = factor(sex_r),
         emconlfreqYes = factor(emconlfreqYes, levels = c("0","1")),       
         lives_onlineYN = factor(lives_onlineYN, levels = c("0","1")),    
         timeexe_r = factor(timeexe_r, levels = c("0","1")),    
         alc30d_YN = factor(alc30d_YN, levels = c("0","1"))          
         )

```

# Univariate EDA
Here for each var, I check it's distribution, summary stats and potential of prediction.

## pmsu_lmh

```{r}

dc %>%
  group_by(pmsu_lmh) %>%
  tally()
xxxx

```


Key Predictors to Study (in this order):

Problematic social media use: 
-  (over pmsu_yn)

Other concerning online stuff:
- emconlfreqYes (too much online comms)
- lives_onlineYN (prefer online vulnerability)

Bulling:
- beenbulliedYes
- cbeenbulliedYes

Youth support system:
- family_support_high (over family_support_hml)
- friends_support_high (over friends_support_hml)
- school_support_high (over teacher_support_high:student_support_high)

Parental support:
- talkfatherYes x talkmotherYes

Habits good/bad:
- timeexe_r (exercise)
- alc30d_YN

Covariates:
- age_r
- sex_r
- IRRELFAS_LMH_r

Other:
- seqno_int
- country_name

Vars to Predict:
- lifesat_low
- mental_complaintsY (or maybe health_complaintsY)



## ROADMAMP
univ eda
see potential of pred for each var
get to know each var








## Univariate Exploration and Preparation  











We now look one by one and will:  
1. Coerce to correct var type for modeling (mostly factors)
2. Note potential of prediction



### seqno_int  
Children id and confirmed it's unique.

```{r}

length(unique(d$seqno_int)) == nrow(d)

```

###$ lifesat_low
13% report having low satisfaction and still ~2% of NAs that need to be fixed. 

```{r, echo=FALSE}

d %>%
  group_by(lifesat_low) %>%
  tally() %>%
  mutate (pct = round(n/sum(n),3))

```

###mental_complaintsY
26% report mental health issues with ~5% of NAs that need to be fixed. 

```{r, echo=FALSE}

d %>%
  group_by(mental_complaintsY) %>%
  tally() %>%
  mutate (pct = round(n/sum(n),3))

```

### country_name
By country, we look at % of children with low lifesat and mental health issues.
xxxxxxxxxxxxxxxxx voy

```{r, echo=FALSE}

d %>%
  group_by(country_name) %>%
  tally() %>%
  mutate (pct = round(n/sum(n),3))

```


###$ pmsu_lmh             <chr> "med", "med", "high", "med", "med", "low", "med", "low", "low", NA, "med", NA, "lo…
#voy, is this right?

```{r}

d <- d %>%
  mutate(pmsu_lmh = factor(pmsu_lmh, 
                           levels = c("Low", "Med", "High")))

d %>%
  filter(!is.na(lifesat_low)
         #& !is.na(pmsu_lmh)
         ) %>%
  group_by(pmsu_lmh,
           lifesat_low) %>%
  summarise(n = n()) %>%
  pivot_wider(values_from = n,
              names_from = lifesat_low,
              names_prefix = "lifesat_low_") %>%
  mutate(n = lifesat_low_1 + lifesat_low_0) %>%
  mutate(lifesat_low_1_pct = round(lifesat_low_1 / n,3))

```


###$ pmsu_yn              <dbl> 0, 0, 1, 0, 0, 0, 0, 0, 0, NA, 0, NA, 0, 0, NA, 0, 0, NA, 0, 0, 0, NA, 0, 0, 0, 0,…

###$ family_support_high  <dbl> 1, 1, 1, 1, 0, 1, 1, NA, 1, 1, NA, NA, 1, 1, 1, 1, 1, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1…

###$ family_support_hml   <chr> "high", "high", "high", "high", "low", "high", "high", NA, "high", "high", NA, NA,…

###$ friends_support_high <dbl> 1, 1, 0, 1, 0, 1, 1, 1, 1, 1, NA, 0, 1, 1, 1, 0, 1, NA, 1, 0, 1, 1, 0, 0, 0, 1, NA…

###$ friends_support_hml  <chr> "high", "high", "med", "high", "high", "high", "high", "high", "high", "high", NA,…

###$ IRRELFAS_LMH_r       <chr> "Med60", NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, "Med60", NA, "Med60", "Med60"…

###$ talkfatherYes        <dbl> 1, 1, 1, 1, 0, 0, 1, NA, NA, 1, NA, NA, 1, NA, 1, 1, 1, NA, 1, 0, 1, 1, 1, 1, 1, 1…

###$ talkmotherYes        <dbl> 1, 1, 1, 1, 1, 1, 1, NA, NA, 1, NA, NA, 1, NA, NA, 1, 1, NA, NA, 0, 1, 1, 1, 1, 1,…

###$ beenbulliedYes       <dbl> 0, 1, 0, 1, 0, 0, 0, 0, 1, 1, 0, 1, 0, 0, 0, 1, 0, 1, 0, 1, 0, 0, 0, 0, 0, 0, 1, 0…

###$ cbeenbulliedYes      <dbl> 1, 1, 1, 0, 0, 0, 0, 0, NA, 0, 0, 0, 0, NA, 0, 0, 0, 1, 0, 1, 0, 0, 0, 0, 0, 0, 0,…

###$ age_r                <dbl> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA…

###$ sex_r  


for later:
keep complete obs only
model like a rusian doll




