---
title: "HBSC: Modeling"
date: "Updated: `r format(Sys.time(), '%Y-%m-%d')`"
output:
  pdf_document: default
  word_document: default
  html_notebook: default
urlcolor: blue
---

```{r project_setup, include=FALSE}

rm(list = ls())

setwd("C:/MisLocalFiles/Github/HBSC/scripts")
project_folder <- "C:/MisLocalFiles/Github/HBSC/"

require(tidyverse)
require(caret)
require(broom)

```

## Data Overview
This data is prepared in hbsc.rmd file. It's dimensions are:

```{r, echo=FALSE}
# ---------------------------------------------------------
# load dataset
# ---------------------------------------------------------

d <- read_csv(paste0(project_folder, "data/processed/dat_2026-08-02.csv"))

dim(d) #244097     17

```
## Univariate Exploration and Preparation  

The list of vars and it's type is:

```{r, echo=FALSE}

glimpse(d)

```
We now look one by one and will:  
1. Coerce to correct var type for modeling (mostly factors)
2. Note potential of prediction

### Coerce first  
For modeling, many vars need to be coerced so I start there before univ EDA.

```{r}

d <- d %>%
  mutate(lifesat_low = factor(lifesat_low, levels = c("0","1")),
         mental_complaintsY = factor(mental_complaintsY, levels = c("0","1")),
         pmsu_lmh = factor(pmsu_lmh, levels = c("low", "med", "high")),
         pmsu_yn = factor(pmsu_yn, levels = c("0", "1")),
         family_support_high = factor(family_support_high, levels = c("0", "1")),
         family_support_hml = factor(family_support_hml, levels = c("low", "med", "high")),
         friends_support_high = factor(friends_support_high, levels = c("0", "1")),
         friends_support_hml = factor(friends_support_hml, levels = c("low", "med", "high")),
         IRRELFAS_LMH_r = factor(IRRELFAS_LMH_r, levels = c("Low20", "Med60", "High20")),
         talkfatherYes = factor(talkfatherYes, levels = c("0", "1")),
         talkmotherYes = factor(talkmotherYes, levels = c("0", "1")),
         beenbulliedYes = factor(beenbulliedYes, levels = c("0", "1")),
         cbeenbulliedYes = factor(cbeenbulliedYes, levels = c("0", "1")),
         age_r = factor(age_r, levels = c("11", "13", "15")),
         sex_r = factor(sex_r)
         )

```

### seqno_int  
Children id and confirmed it's unique.

```{r}

length(unique(d$seqno_int)) == nrow(d)

```

###$ lifesat_low
13% report having low satisfaction and still ~2% of NAs that need to be fixed. 

```{r, echo=FALSE}

d %>%
  group_by(lifesat_low) %>%
  tally() %>%
  mutate (pct = round(n/sum(n),3))

```

###mental_complaintsY
26% report mental health issues with ~5% of NAs that need to be fixed. 

```{r, echo=FALSE}

d %>%
  group_by(mental_complaintsY) %>%
  tally() %>%
  mutate (pct = round(n/sum(n),3))

```

### country_name
By country, we look at % of children with low lifesat and mental health issues.
xxxxxxxxxxxxxxxxx voy

```{r, echo=FALSE}

d %>%
  group_by(country_name) %>%
  tally() %>%
  mutate (pct = round(n/sum(n),3))

```


###$ pmsu_lmh             <chr> "med", "med", "high", "med", "med", "low", "med", "low", "low", NA, "med", NA, "lo…
#voy, is this right?

```{r}

d <- d %>%
  mutate(pmsu_lmh = factor(pmsu_lmh, 
                           levels = c("Low", "Med", "High")))

d %>%
  filter(!is.na(lifesat_low)
         #& !is.na(pmsu_lmh)
         ) %>%
  group_by(pmsu_lmh,
           lifesat_low) %>%
  summarise(n = n()) %>%
  pivot_wider(values_from = n,
              names_from = lifesat_low,
              names_prefix = "lifesat_low_") %>%
  mutate(n = lifesat_low_1 + lifesat_low_0) %>%
  mutate(lifesat_low_1_pct = round(lifesat_low_1 / n,3))

```


###$ pmsu_yn              <dbl> 0, 0, 1, 0, 0, 0, 0, 0, 0, NA, 0, NA, 0, 0, NA, 0, 0, NA, 0, 0, 0, NA, 0, 0, 0, 0,…

###$ family_support_high  <dbl> 1, 1, 1, 1, 0, 1, 1, NA, 1, 1, NA, NA, 1, 1, 1, 1, 1, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1…

###$ family_support_hml   <chr> "high", "high", "high", "high", "low", "high", "high", NA, "high", "high", NA, NA,…

###$ friends_support_high <dbl> 1, 1, 0, 1, 0, 1, 1, 1, 1, 1, NA, 0, 1, 1, 1, 0, 1, NA, 1, 0, 1, 1, 0, 0, 0, 1, NA…

###$ friends_support_hml  <chr> "high", "high", "med", "high", "high", "high", "high", "high", "high", "high", NA,…

###$ IRRELFAS_LMH_r       <chr> "Med60", NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, "Med60", NA, "Med60", "Med60"…

###$ talkfatherYes        <dbl> 1, 1, 1, 1, 0, 0, 1, NA, NA, 1, NA, NA, 1, NA, 1, 1, 1, NA, 1, 0, 1, 1, 1, 1, 1, 1…

###$ talkmotherYes        <dbl> 1, 1, 1, 1, 1, 1, 1, NA, NA, 1, NA, NA, 1, NA, NA, 1, 1, NA, NA, 0, 1, 1, 1, 1, 1,…

###$ beenbulliedYes       <dbl> 0, 1, 0, 1, 0, 0, 0, 0, 1, 1, 0, 1, 0, 0, 0, 1, 0, 1, 0, 1, 0, 0, 0, 0, 0, 0, 1, 0…

###$ cbeenbulliedYes      <dbl> 1, 1, 1, 0, 0, 0, 0, 0, NA, 0, 0, 0, 0, NA, 0, 0, 0, 1, 0, 1, 0, 0, 0, 0, 0, 0, 0,…

###$ age_r                <dbl> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA…

###$ sex_r  


for later:
keep complete obs only
model like a rusian doll





