---
title: 'HBSC: Modeling'
date: "Updated: `r format(Sys.time(), '%Y-%m-%d')`"
output:
  pdf_document: default
  word_document: default
  html_notebook: default
  html_document:
    df_print: paged
urlcolor: blue
---

```{r}

rm(list = ls())

setwd("C:/MisLocalFiles/Github/HBSC/scripts/modeling/")
project_folder <- "C:/MisLocalFiles/Github/HBSC/"

require(tidyverse)
require(caret)
require(broom)
require(kableExtra)
library(knitr)
require(coefplot)
require(sjPlot)
require(shapviz)
require(xgboost)


```

# Data Overview
This data is prepared in hbsc.rmd file. It's dimensions are:

```{r}
# ---------------------------------------------------------
# load dataset
# ---------------------------------------------------------

d <- read_csv(paste0(project_folder, "data/processed/dat_2026-08-16.csv"))

dim(d) #244097     25

```

The list of vars and it's type is:

```{r}

glimpse(d)

```

# Missing Values
Most features have missing values. 

```{r}

summary(d)

```

For now, I will just keep complete observations to simplify my life. 
But as can be seen some countries are very afected. Netherlands is a good option.

```{r}

d <- d %>% 
  mutate(any_na = if_any(everything(), is.na))

d %>%
  group_by(country_name, any_na) %>%
  tally() %>%
  pivot_wider(names_from = any_na,
              names_prefix = "na_",
              values_from = n,
              values_fill = 0) %>%
  mutate(n = na_FALSE + na_TRUE) %>%
  mutate(pct_complete = round(na_FALSE / n,3)) %>%
  select(n, na_FALSE, pct_complete) %>%
  arrange(desc(pct_complete))-> tbl_nas_country

kable(tbl_nas_country)

dc <- d[complete.cases(d),]

summary(dc)

```

# Check if Dataset if balanced based on vars to predict

## lifesat_low
For lifesat_low I like Canada and Netherlands in the extremes. Could be good to compare
if we opt for two countries. For now, I will go with Canada due to big %.

```{r}

dc %>%
  group_by(country_name, lifesat_low) %>%
  tally() %>%
  pivot_wider(names_from = lifesat_low,
              names_prefix = "is_",
              values_from = n,
              values_fill = 0) %>%
  mutate(n = is_0 + is_1) %>%
  mutate(lifesat_low_pct = round(is_1/n,3)) %>%
  select(country_name, n, lifesat_low_pct) %>%
  arrange(desc(lifesat_low_pct)) -> lifesat_low_pct_country

kable(lifesat_low_pct_country)

#mental_complaintsY   

```

## mental_complaintsY
For mental_complaintsY, Netherlads is still good. But Canada is more halfway,
while other countries like Italy are at top. Israel is at top of list on both
so could try that. 

```{r}

dc %>%
  group_by(country_name, mental_complaintsY) %>%
  tally() %>%
  pivot_wider(names_from = mental_complaintsY,
              names_prefix = "is_",
              values_from = n,
              values_fill = 0) %>%
  mutate(n = is_0 + is_1) %>%
  mutate(mental_complaintsY_pct = round(is_1/n,3)) %>%
  select(country_name, n, mental_complaintsY_pct) %>%
  arrange(desc(mental_complaintsY_pct)) -> mental_complaintsY_pct_country

kable(mental_complaintsY_pct_country)

```

Well, ok. I need to make a call for now. So I will pick Italy first since 
it has the highest % of mental health complaints. And if I decide to compare
we can do it with Spain or Netherlands. 

# Coercions of Data Types
For modeling, many vars need to be coerced so I start there before univ EDA.

```{r}

dc <- dc %>%
  mutate(lifesat_low = factor(lifesat_low, levels = c("0","1")),
         mental_complaintsY = factor(mental_complaintsY, levels = c("0","1")),
         health_complaintsY = factor(health_complaintsY, levels = c("0", "1")),
         pmsu_lmh = factor(pmsu_lmh, levels = c("Low", "Med", "High")),
         pmsu_yn = factor(pmsu_yn, levels = c("0", "1")),
         family_support_high = factor(family_support_high, levels = c("0", "1")),
         family_support_hml = factor(family_support_hml, levels = c("Low", "Med", "High")),
         friends_support_high = factor(friends_support_high, levels = c("0", "1")),
         friends_support_hml = factor(friends_support_hml, levels = c("Low", "Med", "High")),
         teacher_support_high = factor(teacher_support_high, levels = c("0", "1")),
         student_support_high = factor(student_support_high, levels = c("0", "1")),
         school_support_high = factor(school_support_high, levels = c("0", "1")),
         IRRELFAS_LMH_r = factor(IRRELFAS_LMH_r, levels = c("Low20", "Med60", "High20")),
         talkfatherYes = factor(talkfatherYes, levels = c("0", "1")),
         talkmotherYes = factor(talkmotherYes, levels = c("0", "1")),
         beenbulliedYes = factor(beenbulliedYes, levels = c("0", "1")),
         cbeenbulliedYes = factor(cbeenbulliedYes, levels = c("0", "1")),
         age_r = factor(age_r, levels = c("11", "13", "15")),
         sex_r = factor(sex_r),
         emconlfreqYes = factor(emconlfreqYes, levels = c("0","1")),       
         lives_onlineYN = factor(lives_onlineYN, levels = c("0","1")),    
         timeexe_r = factor(timeexe_r, levels = c("0","1")),    
         alc30d_YN = factor(alc30d_YN, levels = c("0","1"))          
         )

```


```{r}

italy <- dc %>%
  filter(country_name == "Italy")

summary(italy)

```


# Logical grouping of variables
I group logically in themes the 25 vars as follows:

Vars to Predict:
- lifesat_low
- mental_complaintsY (or maybe health_complaintsY)

Key Predictors to Study (in this order):

Problematic social media use: 
- pmsu_lmh (over pmsu_yn)

Other concerning online stuff:
- emconlfreqYes (too much online comms)
- lives_onlineYN (prefer online vulnerability)

Bulling:
- beenbulliedYes
- cbeenbulliedYes

Youth support system:
- family_support_high (over family_support_hml)
- friends_support_high (over friends_support_hml)
- school_support_high (over teacher_support_high:student_support_high)

Parental support:
- talkfatherYes x talkmotherYes

Habits good/bad:
- timeexe_r (exercise)
- alc30d_YN

Covariates:
- age_r
- sex_r
- IRRELFAS_LMH_r

Other:
- seqno_int
- country_name


# Univariate EDA
Here for each var, I check it's descriptive stats and potential of prediction.

## pmsu_lmh
Has potential based on higher prop of our Y vars based on pmsu.

```{r}

cat ("Frequence counts:")
italy %>%
  group_by(pmsu_lmh) %>%
  tally() %>%
  mutate(prop = n/sum(n))

cat ("Potential of prediction for mental_complaints:")
italy %>%
  group_by(mental_complaintsY, pmsu_lmh) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = mental_complaintsY,
              names_prefix = "mental_complaintsY") %>%
  mutate(n = mental_complaintsY0 + mental_complaintsY1,
         pct_mental_complaintsY1 = mental_complaintsY1 / n)

cat ("Potential of prediction for lifesat_low")
italy %>%
  group_by(lifesat_low, pmsu_lmh) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = lifesat_low,
              names_prefix = "lifesat_lowY") %>%
  mutate(n = lifesat_lowY0 + lifesat_lowY1,
         pct_lifesat_lowY1 = lifesat_lowY1 / n)

```

## emconlfreqYes (too much online comms)
Has potential for mental issues but doesn't for lifesatlow.

```{r}

cat ("Frequence counts:")
italy %>%
  group_by(emconlfreqYes) %>%
  tally() %>%
  mutate(prop = n/sum(n))

cat ("Potential of prediction for mental_complaints:")
italy %>%
  group_by(mental_complaintsY, emconlfreqYes) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = mental_complaintsY,
              names_prefix = "mental_complaintsY") %>%
  mutate(n = mental_complaintsY0 + mental_complaintsY1,
         pct_mental_complaintsY1 = mental_complaintsY1 / n)

cat ("Potential of prediction for lifesat_low")
italy %>%
  group_by(lifesat_low, emconlfreqYes) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = lifesat_low,
              names_prefix = "lifesat_lowY") %>%
  mutate(n = lifesat_lowY0 + lifesat_lowY1,
         pct_lifesat_lowY1 = lifesat_lowY1 / n)

```

## lives_onlineYN (prefer online vulnerability)
Has potential for both vars. 

```{r}

cat ("Frequence counts:")
italy %>%
  group_by(lives_onlineYN) %>%
  tally() %>%
  mutate(prop = n/sum(n))

cat ("Potential of prediction for mental_complaints:")
italy %>%
  group_by(mental_complaintsY, lives_onlineYN) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = mental_complaintsY,
              names_prefix = "mental_complaintsY") %>%
  mutate(n = mental_complaintsY0 + mental_complaintsY1,
         pct_mental_complaintsY1 = mental_complaintsY1 / n)

cat ("Potential of prediction for lifesat_low")
italy %>%
  group_by(lifesat_low, lives_onlineYN) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = lifesat_low,
              names_prefix = "lifesat_lowY") %>%
  mutate(n = lifesat_lowY0 + lifesat_lowY1,
         pct_lifesat_lowY1 = lifesat_lowY1 / n)

```

## beenbulliedYes
Small % has been bullied and has strong potential for both vars. 

```{r}

cat ("Frequence counts:")
italy %>%
  group_by(beenbulliedYes) %>%
  tally() %>%
  mutate(prop = n/sum(n))

cat ("Potential of prediction for mental_complaints:")
italy %>%
  group_by(mental_complaintsY, beenbulliedYes) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = mental_complaintsY,
              names_prefix = "mental_complaintsY") %>%
  mutate(n = mental_complaintsY0 + mental_complaintsY1,
         pct_mental_complaintsY1 = mental_complaintsY1 / n)

cat ("Potential of prediction for lifesat_low")
italy %>%
  group_by(lifesat_low, beenbulliedYes) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = lifesat_low,
              names_prefix = "lifesat_lowY") %>%
  mutate(n = lifesat_lowY0 + lifesat_lowY1,
         pct_lifesat_lowY1 = lifesat_lowY1 / n)

```

## cbeenbulliedYes
Same as previous bullying metric. 

```{r}

cat ("Frequence counts:")
italy %>%
  group_by(cbeenbulliedYes) %>%
  tally() %>%
  mutate(prop = n/sum(n))

cat ("Potential of prediction for mental_complaints:")
italy %>%
  group_by(mental_complaintsY, cbeenbulliedYes) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = mental_complaintsY,
              names_prefix = "mental_complaintsY") %>%
  mutate(n = mental_complaintsY0 + mental_complaintsY1,
         pct_mental_complaintsY1 = mental_complaintsY1 / n)

cat ("Potential of prediction for lifesat_low")
italy %>%
  group_by(lifesat_low, cbeenbulliedYes) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = lifesat_low,
              names_prefix = "lifesat_lowY") %>%
  mutate(n = lifesat_lowY0 + lifesat_lowY1,
         pct_lifesat_lowY1 = lifesat_lowY1 / n)

```

## family_support_high (over family_support_hml)
Highest % have support and that certainly helps lower the predictions. 

```{r}

cat ("Frequence counts:")
italy %>%
  group_by(family_support_high) %>%
  tally() %>%
  mutate(prop = n/sum(n))

cat ("Potential of prediction for mental_complaints:")
italy %>%
  group_by(mental_complaintsY, family_support_high) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = mental_complaintsY,
              names_prefix = "mental_complaintsY") %>%
  mutate(n = mental_complaintsY0 + mental_complaintsY1,
         pct_mental_complaintsY1 = mental_complaintsY1 / n)

cat ("Potential of prediction for lifesat_low")
italy %>%
  group_by(lifesat_low, family_support_high) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = lifesat_low,
              names_prefix = "lifesat_lowY") %>%
  mutate(n = lifesat_lowY0 + lifesat_lowY1,
         pct_lifesat_lowY1 = lifesat_lowY1 / n)

```

## friends_support_high (over friends_support_hml)
Same as above. 

```{r}

cat ("Frequence counts:")
italy %>%
  group_by(friends_support_high) %>%
  tally() %>%
  mutate(prop = n/sum(n))

cat ("Potential of prediction for mental_complaints:")
italy %>%
  group_by(mental_complaintsY, friends_support_high) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = mental_complaintsY,
              names_prefix = "mental_complaintsY") %>%
  mutate(n = mental_complaintsY0 + mental_complaintsY1,
         pct_mental_complaintsY1 = mental_complaintsY1 / n)

cat ("Potential of prediction for lifesat_low")
italy %>%
  group_by(lifesat_low, friends_support_high) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = lifesat_low,
              names_prefix = "lifesat_lowY") %>%
  mutate(n = lifesat_lowY0 + lifesat_lowY1,
         pct_lifesat_lowY1 = lifesat_lowY1 / n)

```

## school_support_high (over teacher_support_high:student_support_high)
More balanced but having support helps.

```{r}

cat ("Frequence counts:")
italy %>%
  group_by(school_support_high) %>%
  tally() %>%
  mutate(prop = n/sum(n))

cat ("Potential of prediction for mental_complaints:")
italy %>%
  group_by(mental_complaintsY, school_support_high) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = mental_complaintsY,
              names_prefix = "mental_complaintsY") %>%
  mutate(n = mental_complaintsY0 + mental_complaintsY1,
         pct_mental_complaintsY1 = mental_complaintsY1 / n)

cat ("Potential of prediction for lifesat_low")
italy %>%
  group_by(lifesat_low, school_support_high) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = lifesat_low,
              names_prefix = "lifesat_lowY") %>%
  mutate(n = lifesat_lowY0 + lifesat_lowY1,
         pct_lifesat_lowY1 = lifesat_lowY1 / n)

```


## timeexe_r (exercise)
Most exercise and seems to help a bit, not much but some. 

```{r}

cat ("Frequence counts:")
italy %>%
  group_by(timeexe_r) %>%
  tally() %>%
  mutate(prop = n/sum(n))

cat ("Potential of prediction for mental_complaints:")
italy %>%
  group_by(mental_complaintsY, timeexe_r) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = mental_complaintsY,
              names_prefix = "mental_complaintsY") %>%
  mutate(n = mental_complaintsY0 + mental_complaintsY1,
         pct_mental_complaintsY1 = mental_complaintsY1 / n)

cat ("Potential of prediction for lifesat_low")
italy %>%
  group_by(lifesat_low, timeexe_r) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = lifesat_low,
              names_prefix = "lifesat_lowY") %>%
  mutate(n = lifesat_lowY0 + lifesat_lowY1,
         pct_lifesat_lowY1 = lifesat_lowY1 / n)

```

## alc30d_YN
1 in 4 drink and that is correlated with more issues. 

```{r}

cat ("Frequence counts:")
italy %>%
  group_by(alc30d_YN) %>%
  tally() %>%
  mutate(prop = n/sum(n))

cat ("Potential of prediction for mental_complaints:")
italy %>%
  group_by(mental_complaintsY, alc30d_YN) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = mental_complaintsY,
              names_prefix = "mental_complaintsY") %>%
  mutate(n = mental_complaintsY0 + mental_complaintsY1,
         pct_mental_complaintsY1 = mental_complaintsY1 / n)

cat ("Potential of prediction for lifesat_low")
italy %>%
  group_by(lifesat_low, alc30d_YN) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = lifesat_low,
              names_prefix = "lifesat_lowY") %>%
  mutate(n = lifesat_lowY0 + lifesat_lowY1,
         pct_lifesat_lowY1 = lifesat_lowY1 / n)

```

## age_r
Things to get worse as kids grow up.

```{r}

cat ("Frequence counts:")
italy %>%
  group_by(age_r) %>%
  tally() %>%
  mutate(prop = n/sum(n))

cat ("Potential of prediction for mental_complaints:")
italy %>%
  group_by(mental_complaintsY, age_r) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = mental_complaintsY,
              names_prefix = "mental_complaintsY") %>%
  mutate(n = mental_complaintsY0 + mental_complaintsY1,
         pct_mental_complaintsY1 = mental_complaintsY1 / n)

cat ("Potential of prediction for lifesat_low")
italy %>%
  group_by(lifesat_low, age_r) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = lifesat_low,
              names_prefix = "lifesat_lowY") %>%
  mutate(n = lifesat_lowY0 + lifesat_lowY1,
         pct_lifesat_lowY1 = lifesat_lowY1 / n)

```

## sex_r
Balanced dataset with girls having a harder time. 

```{r}

cat ("Frequence counts:")
italy %>%
  group_by(sex_r) %>%
  tally() %>%
  mutate(prop = n/sum(n))

cat ("Potential of prediction for mental_complaints:")
italy %>%
  group_by(mental_complaintsY, sex_r) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = mental_complaintsY,
              names_prefix = "mental_complaintsY") %>%
  mutate(n = mental_complaintsY0 + mental_complaintsY1,
         pct_mental_complaintsY1 = mental_complaintsY1 / n)

cat ("Potential of prediction for lifesat_low")
italy %>%
  group_by(lifesat_low, sex_r) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = lifesat_low,
              names_prefix = "lifesat_lowY") %>%
  mutate(n = lifesat_lowY0 + lifesat_lowY1,
         pct_lifesat_lowY1 = lifesat_lowY1 / n)

```

## IRRELFAS_LMH_r
Yes, money helps.

```{r}

cat ("Frequence counts:")
italy %>%
  group_by(IRRELFAS_LMH_r) %>%
  tally() %>%
  mutate(prop = n/sum(n))

cat ("Potential of prediction for mental_complaints:")
italy %>%
  group_by(mental_complaintsY, IRRELFAS_LMH_r) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = mental_complaintsY,
              names_prefix = "mental_complaintsY") %>%
  mutate(n = mental_complaintsY0 + mental_complaintsY1,
         pct_mental_complaintsY1 = mental_complaintsY1 / n)

cat ("Potential of prediction for lifesat_low")
italy %>%
  group_by(lifesat_low, IRRELFAS_LMH_r) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = lifesat_low,
              names_prefix = "lifesat_lowY") %>%
  mutate(n = lifesat_lowY0 + lifesat_lowY1,
         pct_lifesat_lowY1 = lifesat_lowY1 / n)

```

## talkfatherYes x talkmotherYes
Talking to father really helps.

```{r}

cat ("Frequence counts:")
italy %>%
  group_by(talkfatherYes) %>%
  tally() %>%
  mutate(prop = n/sum(n))

cat ("Potential of prediction for mental_complaints:")
italy %>%
  group_by(mental_complaintsY, talkfatherYes) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = mental_complaintsY,
              names_prefix = "mental_complaintsY") %>%
  mutate(n = mental_complaintsY0 + mental_complaintsY1,
         pct_mental_complaintsY1 = mental_complaintsY1 / n)

cat ("Potential of prediction for lifesat_low")
italy %>%
  group_by(lifesat_low, talkfatherYes) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = lifesat_low,
              names_prefix = "lifesat_lowY") %>%
  mutate(n = lifesat_lowY0 + lifesat_lowY1,
         pct_lifesat_lowY1 = lifesat_lowY1 / n)

```

But it seems talking to a mother even helps more!

```{r}

cat ("Frequence counts:")
italy %>%
  group_by(talkmotherYes) %>%
  tally() %>%
  mutate(prop = n/sum(n))

cat ("Potential of prediction for mental_complaints:")
italy %>%
  group_by(mental_complaintsY, talkmotherYes) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = mental_complaintsY,
              names_prefix = "mental_complaintsY") %>%
  mutate(n = mental_complaintsY0 + mental_complaintsY1,
         pct_mental_complaintsY1 = mental_complaintsY1 / n)

cat ("Potential of prediction for lifesat_low")
italy %>%
  group_by(lifesat_low, talkmotherYes) %>%
  tally() %>%
  pivot_wider(values_from = n,
              names_from = lifesat_low,
              names_prefix = "lifesat_lowY") %>%
  mutate(n = lifesat_lowY0 + lifesat_lowY1,
         pct_lifesat_lowY1 = lifesat_lowY1 / n)

```

## seqno_int

```{r}

# has to be true as identifier
length(unique(italy$seqno_int)) == nrow(italy)

```

# Logistic Regression
Fit with all vars and use forward variable selection.

## For lifesat_low:

```{r}

# Fit GML model using forward subset selection
reg0 <- glm(lifesat_low ~ pmsu_lmh, data = italy, family = binomial) 
reg1 <- glm(lifesat_low ~  pmsu_lmh +
              emconlfreqYes +
              lives_onlineYN +
              family_support_high +
              friends_support_high +
              school_support_high +
              IRRELFAS_LMH_r +
              talkfatherYes +
              talkmotherYes +
              beenbulliedYes +
              cbeenbulliedYes +
              timeexe_r +
              alc30d_YN +
              age_r +
              sex_r, 
            data = italy, 
            family = binomial)

summary(reg0)
summary(reg1)

best <- step(reg0, scope = formula(reg1), 
             direction = "forward", k = 2) #k=2 is AIC
summary(best)
best$anova

coefplot(best, intercept = FALSE, title = "Log-Odds Coefficients")
mytidy <- tidy(best, exponentiate = TRUE, conf.int = TRUE)

ggplot(mytidy, aes(x = term, y = estimate)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                width = 0.15, color = "gray40") +
  geom_point(size = 3, color = "blue") +
  coord_flip() +
  labs(
    title = "Coefficient Estimates with 95% Confidence Intervals",
    x = "",
    y = "Estimate"
  ) +
  theme_minimal(base_size = 14)



```

## For mental_complaintsY:

```{r}

# Fit GML model using forward subset selection
reg0 <- glm(mental_complaintsY ~ pmsu_lmh, data = italy, family = binomial) 
reg1 <- glm(mental_complaintsY ~  pmsu_lmh +
              emconlfreqYes +
              lives_onlineYN +
              family_support_high +
              friends_support_high +
              school_support_high +
              IRRELFAS_LMH_r +
              talkfatherYes +
              talkmotherYes +
              beenbulliedYes +
              cbeenbulliedYes +
              timeexe_r +
              alc30d_YN +
              age_r +
              sex_r, 
            data = italy, 
            family = binomial)

summary(reg0)
summary(reg1)

best <- step(reg0, scope = formula(reg1), 
             direction = "forward", k = 2) #k=2 is AIC
summary(best)
best$anova

coefplot(best, intercept = FALSE, title = "Log-Odds Coefficients")

mytidy <- tidy(best, exponentiate = TRUE, conf.int = TRUE)
ggplot(mytidy, aes(x = term, y = estimate)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                width = 0.15, color = "gray40") +
  geom_point(size = 3, color = "blue") +
  coord_flip() +
  labs(
    #title = "Coefficient Estimates with 95% Confidence Intervals",
    x = "",
    y = "Estimate"
  ) +
  theme_minimal(base_size = 14)



```

## The first SHAP


```{r}

italy2 <- italy

italy2$mental_complaintsY_num <- as.numeric(italy2$mental_complaintsY)
#italy2$pmsu_yn_num <- as.numeric(italy2$pmsu_yn)
#italy2$beenbulliedYes_num <- as.numeric(italy2$beenbulliedYes)

set.seed(3653)
#X <- italy2[c("pmsu_yn_num", "beenbulliedYes_num")]
X <- italy2[c("pmsu_yn", "beenbulliedYes")]
dtrain <- xgb.DMatrix(data.matrix(X), label = italy2$mental_complaintsY_num)

fit <- xgb.train(
  params = list(learning_rate = 0.1, objective = "reg:squarederror"), 
  data = dtrain,
  nrounds = 65L
)

# Explanation dataset
X_small <- X[sample(nrow(X), 1000L), ]

shp <- shapviz(fit, X_pred = data.matrix(X_small), X = X_small)

sv_waterfall(shp, row_id = 1)

sv_force(shp, row_id = 10)
sv_force(shp, row_id = 1)

sv_importance(shp)
sv_importance(shp,
              kind = "beeswarm")

sv_importance(shp, kind = "both", alpha = 0.2, width = 0.2)

sv_dependence(shp, v = "pmsu_yn", "auto")

```

chi2 test hacer


https://www.r-bloggers.com/2023/09/decisions-at-the-door-understanding-logistic-regression-with-tidymodels/
https://www.tidymodels.org/
https://www.tmwr.org/
https://medium.com/data-science/shap-for-binary-and-multiclass-target-variables-ff2f43de0cf4
https://benhay.es/posts/interpreting-ml-shap/





