---
title: "HBSC"
date: "Updated: `r format(Sys.time(), '%Y-%m-%d')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
urlcolor: blue
---

```{r setup, include=FALSE}

rm(list = ls())

require(tidyverse)
library(knitr)
require(stringr)
library(scales)
library(corrplot)
library(psych)



#prj_fldr <- "C:/MisLocalFiles/Github/HBSC/"
prj_fldr <- "C:/Github/HBSC/"
#setwd(prj_fldr)
#knitr::opts_knit\$set(root.dir = "C:/MisLocalFiles/Github/HBSC")

```

```{r dat_read, echo=FALSE}

dat_fldr <- paste0(prj_fldr,"data/")

hbsc2018 <- read_delim(paste0(dat_fldr, "raw/HBSC2018OAed1.1.csv"),
                       delim = ";",
                       na = c("", " "),
                       show_col_types = FALSE
                       )

```

# Data Overview

It's a huge file:

```{r, echo=FALSE}

dim(hbsc2018)

```

With these columns:

```{r, echo=FALSE}

# get column names
cols <- colnames(hbsc2018)
#cols <- strsplit(cols, ";")[[1]]
#cols <- trimws(cols)

n <- length(cols)
mat <- matrix(cols, ncol = 5, byrow = TRUE)

kable(mat)

```

So after selecting and grouping variables, create dat with :

```{r dat_create, echo=FALSE}

var_groups <- list(
  "Individual Information" = c(
    "seqno_int", "countryno", "weight", "agecat", "sex",
    "MBMI", "IRFAS", "IRRELFAS_LMH", "oweight_who",
    "bodyweight", "bodyheight"
  ),
  "Health Issues and Life Satisfaction" = c(
    "health", "lifesat", "headache", "stomachache", "backache",
    "feellow", "irritable", "nervous", "sleepdificulty", "dizzy"
  ),
  "Problematic Social Media Use (PMSU)" = c(
    "emcsocmed1", "emcsocmed2", "emcsocmed3",
    "emcsocmed4", "emcsocmed5", "emcsocmed6",
    "emcsocmed7", "emcsocmed8", "emcsocmed9"
  ),
  "Bullying" = c(
    "beenbullied", "cbeenbullied"
  ),
  "Communicates Online More Than Real Life" = c(
    "emconlfreq1", "emconlfreq2", "emconlfreq3", "emconlfreq4",
    "emconlpref1", "emconlpref2", "emconlpref3"
  ),
  "Family Support" = c(
    "famhelp", "famsup", "famtalk", "famdec"
  ),
  "Friends Support" = c(
    "friendhelp", "friendcounton", "friendshare", "friendtalk"
  ),
  "School Support" = c(
    "likeschool", "schoolpressure",
    "studtogether", "studhelpful", "studaccept",
    "teacheraccept", "teachercare", "teachertrust"
  ),
  "Communication with Parents" = c(
    "talkfather", "talkmother"
  ),
  "Good/Bad Habits" = c(
    "physact60", "timeexe", "alc30d_2"
  )
)

# create dat
cols_to_keep <- unlist(var_groups, use.names = FALSE)
dat <- hbsc2018 %>%
  select(all_of(cols_to_keep))

# prettt print vars
 for (group_name in names(var_groups)) {
   cat("##", group_name, "\n")
 
   cat(
     paste0(
       "* `", var_groups[[group_name]], "`",
       collapse = "\n"
     ),
     "\n\n"
   )
 }

```

The number of missing values is relevant and need to address later:

```{r nas, echo=FALSE}
na_cnts <- data.frame(
  variable = names(dat),
  na_count = sapply(dat, function(x) sum(is.na(x)))
)

row.names(na_cnts) <- NULL

na_cnts <- na_cnts %>%
  #mutate(na_pct = round(100 * na_count / nrow(dat),0)) %>%
  mutate(na_pct = na_count / nrow(dat)) %>% 
  mutate(na_pct = percent(na_pct, accuracy = 1)) %>%
  arrange(desc(na_count))

print(na_cnts)
```

Now I explore all variables by group and feature engineer as appropriate.

# [1] "Individual Information"                 

11 vars here. So do 1x1.

```{r ii, echo=FALSE}

vars_indivinfo <- var_groups[["Individual Information"]]
vars_indivinfo

```

seqno_int is just an identifier

```{r, include=FALSE}

length(table(dat$seqno_int)) == nrow(dat)

```
countryno is recoded to country_name and this is the distribution.

```{r country, echo=FALSE}

recode_map <- c(
  "8000"   = "Albania",
  "31000"  = "Azerbaijan",
  "40000"  = "Austria",
  "51000"  = "Armenia",
  "56001"  = "Belgium (Flemish)",
  "56002"  = "Belgium (French)",
  "100000" = "Bulgaria",
  "124000" = "Canada",
  "191000" = "Croatia",
  "203000" = "Czech Republic",
  "208000" = "Denmark",
  "233000" = "Estonia",
  "246000" = "Finland",
  "250000" = "France",
  "268000" = "Georgia",
  "276000" = "Germany",
  "300000" = "Greece",
  "304000" = "Greenland",
  "348000" = "Hungary",
  "352000" = "Iceland",
  "372000" = "Ireland",
  "376000" = "Israel",
  "380000" = "Italy",
  "398000" = "Kazakhstan",
  "428000" = "Latvia",
  "440000" = "Lithuania",
  "442000" = "Luxembourg",
  "470000" = "Malta",
  "498000" = "Republic of Moldova",
  "528000" = "Netherlands",
  "578000" = "Norway",
  "616000" = "Poland",
  "620000" = "Portugal",
  "642000" = "Romania",
  "643000" = "Russia",
  "688000" = "Serbia",
  "703000" = "Slovakia",
  "705000" = "Slovenia",
  "724000" = "Spain",
  "752000" = "Sweden",
  "756000" = "Switzerland",
  "792000" = "Turkey",
  "804000" = "Ukraine",
  "807000" = "Macedonia",
  "826001" = "England",
  "826002" = "Scotland",
  "826003" = "Wales",
  "826004" = "Northern Ireland",
  "840000" = "USA"
)

dat$country_name <- recode_map[as.character(dat$countryno)]

tbl_country <- dat %>%
  group_by(country_name) %>%
  summarise(n = n(),
            .groups = "drop") %>%
  #mutate(pct = round(100 * n / sum(n),1)) %>%
  mutate(pct = n / sum(n)) %>%  
  mutate(pct = percent(pct, accuracy = 1)) %>% 
  arrange(country_name) 

kable(tbl_country,
      caption = "#/% of observations by country")

dat <- dat %>% 
  select(-countryno)

```

weight is unrelated so ignore/remove TBC?

```{r weig, include=FALSE}

dat$weight

```

bodyheight seems to be cm and range from 120 to 200 so will use. Impute NAs?

```{r, echo=FALSE}

summary(dat$bodyheight)

ggplot(dat, aes(x = bodyheight)) +
  geom_histogram() +
  #facet_grid(sex ~ .) +
  theme_minimal()

```

bodyweight seems to be kg? and range from 20 to 150 so will use maybe. Impute NAs?

```{r, echo=FALSE}

summary(dat$bodyweight)

ggplot(dat, aes(x = bodyweight)) +
  geom_histogram() +
  #facet_grid(sex ~ .) +
  theme_minimal()

```


MBMI TBC huge odd range of #s.

```{r, echo=FALSE}

summary(dat$MBMI)

ggplot(dat, aes(x = MBMI)) +
  geom_histogram() +
  #facet_grid(sex ~ .) +
  theme_minimal()

```

oweight_who seems to be binary??? TBC. Not sure if helpful to add.

```{r ow, echo=FALSE}

table(dat$oweight_who, useNA = "always")

```

agecat has 3 values and will leave as is: 1 ~ 11, 2 ~ 13, 3 ~ 15.

```{r age, echo=FALSE}

table(dat$agecat, useNA = "always")

```

sex is binary with 1 (boy), 2 (girl) so will leave as is.

```{r sex, echo=FALSE}

table(dat$sex, useNA = "always")

```

IRRELFAS_LMH is relative family affluence with 1 Low20, 2 Mid60, 3 High20. Leave as is.

IRFAS is discrete [0,13]. TBC what it means and rel with LMH cause it's not mut exclusive. 

```{r fas, echo=FALSE}

#table(dat$IRRELFAS_LMH, useNA = "always")

#table(dat$IRFAS)

table(dat$IRFAS,
      dat$IRRELFAS_LMH,
      useNA = "always")

```



# [3] "Problematic Social Media Use (PMSU)"  
9 binary variables are recoded to 0/1.

```{r pmsu_recode, echo=FALSE}

vars_pmsu <- var_groups[["Problematic Social Media Use (PMSU)"]]
vars_pmsu_r <- paste0(vars_pmsu, "_r")
  
# recode to 0/1 from 1/2
recode_emcsocmed <- function(old_col){
  new_col <- case_when(
    old_col == 1 ~ 0,
    old_col == 2 ~ 1,
    TRUE ~ NA #99?
  )
  return(new_col)
}

dat <- dat %>%
  mutate(emcsocmed1_r = recode_emcsocmed(emcsocmed1),
         emcsocmed2_r = recode_emcsocmed(emcsocmed2),
         emcsocmed3_r = recode_emcsocmed(emcsocmed3),
         emcsocmed4_r = recode_emcsocmed(emcsocmed4),
         emcsocmed5_r = recode_emcsocmed(emcsocmed5),
         emcsocmed6_r = recode_emcsocmed(emcsocmed6),
         emcsocmed7_r = recode_emcsocmed(emcsocmed7),
         emcsocmed8_r = recode_emcsocmed(emcsocmed8),
         emcsocmed9_r = recode_emcsocmed(emcsocmed9)
         ) 

# and look like pprrint:
questions <- data.frame(
  item = c(
    "emcsocmed1","emcsocmed2","emcsocmed3","emcsocmed4",
    "emcsocmed5","emcsocmed6","emcsocmed7","emcsocmed8","emcsocmed9"
  ),
  topic = c(
    "Social media: I can't think of anything else",
    "Social media: Dissatisfied cause want to spend more time",
    "Social media: Felt bad cause want to play",
    "Social media: Failed to spend less time",
    "Social media: Neglected other activities",
    "Social media: Had arguments because of use",
    "Social media: Lied about amount of time spent",
    "Social media: Use it to escape from negative feelings",
    "Social media: Conflict with family because of use"
  )
)

kable(questions) #, caption = "Social Media Use Items")

```

And they are correlated.

```{r corr2,echo=FALSE}

tetra_mat <- tetrachoric(dat[vars_pmsu_r])$rho
tetra_mat <- round(tetra_mat, 2)

corrplot::corrplot(
  tetra_mat,
  method = "color",
  type = "upper",
  tl.col = "black",
  addCoef.col = "black"
)

```

These 9 vars are sumed and then binned to H/M/L where H(6-9), M(2-5) and L(0-1). Resulting in:

```{r pmsu_recode2, echo=FALSE}

#sum and bin
dat <- dat %>%
  # should go from [0,9]
  mutate(pmsu_sum = rowSums(dat[,vars_pmsu_r])) %>%
  # 3 groups
  mutate(
    pmsu_lmh = case_when(
      between(pmsu_sum,6,9) ~ 3, #"High",
      between(pmsu_sum,2,5) ~ 2, #"Med",
      between(pmsu_sum,0,1) ~ 1, #"Low",
      TRUE ~ NA
    )
  )

#cat("## Table for emcsocmed_sum x pmsu_lmh")
kable(table(dat$pmsu_sum,
      dat$pmsu_lmh,
      useNA = "always")
)

# remove original vars
dat <-  dat %>%
  select(-all_of(vars_pmsu))

```


# [4] "Bullying"                               

There are 2 variables that measure if kid has been bullied. a) beenbullied: Been bullied past months
and b) cbeenbullied: Been cyber bullied. Answers range from 1 (Haven't) to 5 (Several times a week).  

I will derive 2 to keep: cbeenbullied as y/n and bulliedsum as sum of both.

```{r bully, echo=FALSE}

dat <- dat %>%
  mutate(cbullied = 
           case_when(cbeenbullied == 1 ~ 0,
                     between(cbeenbullied, 2, 5) ~ 1,
                     TRUE ~ NA),
         bulliedsum = beenbullied + cbeenbullied
           )

cat("cbullied recoding:")
table(dat$cbeenbullied,
      dat$cbullied,
      useNA = "always")

bullydf <- data.frame(table(dat$bulliedsum, useNA = "always")) %>%
  rename(score = Var1,
         n = Freq) %>%
  mutate(pct = n / sum(n),
         pct = percent(pct, accuracy = 1))

cat("bullying score: (low is better)")
kable(bullydf)

```


# [6] "Family Support" 
4 questions going from 1 Very strongly disagree to 7 Very strongly agree.

```{r famprnt, echo=FALSE}

vars_famsup <- var_groups[["Family Support"]]
vars_famsup

family_items <- data.frame(
  item = c(
    "famhelp",
    "famsup",
    "famtalk",
    "famclose"
  ),
  topic = c(
    "Family tries to help",
    "Get emotional help from family",
    "Talk about problems with family",
    "Feel close to family"
  )
)

kable(family_items) #, caption = "Family Support Items")
```

And they are correlated.

```{r famcorr2,echo=FALSE}

#tetra_mat <- tetrachoric(dat[vars_famsup])$rho
cor_famsup <- cor(dat[vars_famsup],
    use = "pairwise.complete.obs")

#tetra_mat <- round(tetra_mat, 2)

corrplot::corrplot(
  cor_famsup,
  method = "color",
  type = "upper",
  tl.col = "black",
  addCoef.col = "black"
)

```

Summing them results in [4,28] and then grouped as H/M/L as L 4-11, M 12-19, H 20-28. Resulting in:

```{r fam_rec, echo=FALSE}

dat <- dat %>%
   mutate(famsup_sum = rowSums(dat[,vars_famsup])
          ) %>%
   mutate(famsup_hml = case_when(
            famsup_sum >= 4 & famsup_sum <= 11 ~ 1, #"Low",
            famsup_sum >= 12 & famsup_sum <= 19 ~ 2, #"Med",
            famsup_sum >= 20 & famsup_sum <= 28 ~ 3, #"High",
            TRUE ~ NA))

tbl_famsup <- table(dat$famsup_sum,
                    dat$famsup_hml,
                    useNA = "always"
                    )

kable(tbl_famsup)

```


# [7] "Friends Support"   
4 questions going from 1 Very strongly disagree to 7 Very strongly agree.

```{r friprnt, echo=FALSE}

vars_frisup <- var_groups[["Friends Support"]]
vars_frisup

friend_items <- data.frame(
  item = c(
    "friendhelp",
    "friendcounton",
    "friendshare",
    "friendtalk"
  ),
  topic = c(
    "My friends really try to help me",
    "I can count on my friends when things go wrong",
    "I have friends with whom I can share my joys and sorrows",
    "I can talk about my problems with my friends"
  )
)

kable(friend_items)

```

And they are correlated.

```{r frisup_r,echo=FALSE}

#tetra_mat <- tetrachoric(dat[vars_famsup])$rho
cor_frisup <- cor(dat[vars_frisup],
    use = "pairwise.complete.obs")

#tetra_mat <- round(tetra_mat, 2)

corrplot::corrplot(
  cor_frisup,
  method = "color",
  type = "upper",
  tl.col = "black",
  addCoef.col = "black"
)

```

Summing them results in [4,28] and then grouped as H/M/L as L 4-11, M 12-19, H 20-28. Resulting in:

```{r fri_rec, echo=FALSE}

dat <- dat %>%
   mutate(frisup_sum = rowSums(dat[,vars_frisup])
          ) %>%
   mutate(frisup_hml = case_when(
            frisup_sum >= 4 & frisup_sum <= 11 ~ 1, #"Low",
            frisup_sum >= 12 & frisup_sum <= 19 ~ 2, #"Med",
            frisup_sum >= 20 & frisup_sum <= 28 ~ 3, #"High",
            TRUE ~ NA))

tbl_frisup <- table(dat$frisup_sum,
                    dat$frisup_hml,
                    useNA = "always"
                    )

kable(tbl_frisup)

```

# [9] "Communication with Parents"

Two vars with answers from 1 (Very easy) to 5 (Don't have or see). With this counts and proportions in data.  

```{r commp, echo=FALSE}

vars_commparents <- var_groups[["Communication with Parents"]]
vars_commparents

table(dat[vars_commparents], useNA = "always")
round(table(dat[vars_commparents], useNA = "always") / nrow(dat),3)

```

With some correlation between them (I expected more).

```{r cpcorr, echo=FALSE}

corrmat <- cor(dat[vars_commparents], 
               use = "pairwise.complete.obs")

```

So I will dichotomize like HBSC and keep sum of both as well for SHAP. These are dichotomized for each parent as Yes if [1,2] and No if [3-5].  

```{r, include=FALSE}

dat <- dat %>%
  mutate(talkf = 
           case_when(talkfather %in% c(1,2) ~ 1,
                     talkfather %in% c(3,4,5) ~ 0,
                     TRUE ~ NA),
         talkm = 
           case_when(talkmother %in% c(1,2) ~ 1,
                     talkmother %in% c(3,4,5) ~ 0,
                     TRUE ~ NA),
         talkscore = talkfather + talkmother
         )

```

Where new vars look like this for dichotom values:

```{r prcom, echo=FALSE}

cat("Father:")
table(dat$talkfather,
      dat$talkf,
      useNA = "always")

cat("Mother:")
table(dat$talkfather,
      dat$talkf,
      useNA = "always")

```

And the talk score looks like this. Will keep for now and orig vars as well.

```{r prsum, echo=FALSE}

#summary(dat$talkscore)

ggplot(dat, aes(x = factor(talkscore))) +
  geom_bar() +
  #scale_x_discrete() +
  theme_minimal()

```

voy
[2] "Health Issues and Life Satisfaction"    
[5] "Communicates Online More Than Real Life"
[8] "School Support"                         
[10] "Good/Bad Habits"









