---
title: "HBSC"
date: "Updated: `r format(Sys.time(), '%Y-%m-%d')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r projset, echo=FALSE}

rm(list = ls())

require(tidyverse)
require(tidymodels)
library(knitr)
require(stringr)
library(scales)
library(corrplot)
library(psych)
require(rsample)

prj_fldr <- "C:/MisLocalFiles/Github/HBSC/"
#prj_fldr <- "C:/Github/HBSC/"

```


```{r dat_read, echo=FALSE}

dat_fldr <- paste0(prj_fldr,"data/")

# hbsc2018 <- read_delim(paste0(dat_fldr, "raw/HBSC2018OAed1.1.csv"),
#                        delim = ";",
#                        na = c("", " "),
#                        show_col_types = FALSE
#                        )

# the european way with ; and ,
hbsc2018 <- read_csv2(paste0(dat_fldr, "raw/HBSC2018OAed1.1.csv"))

```

# Data Overview

It's a huge file:

```{r, echo=FALSE}

dim(hbsc2018)

```

With these columns:

```{r, echo=FALSE}

# get column names
cols <- colnames(hbsc2018)
#cols <- strsplit(cols, ";")[[1]]
#cols <- trimws(cols)

n <- length(cols)
mat <- matrix(cols, ncol = 5, byrow = TRUE)

kable(mat)

```

So after selecting and grouping variables, create dat with :

```{r dat_create, echo=FALSE}

var_groups <- list(
  "Individual Information" = c(
    "seqno_int", "countryno",  "agecat", "sex",
    "MBMI", "IRFAS", "IRRELFAS_LMH", "IOTF4",
    "bodyweight", "bodyheight"
  ),
  "Health Issues and Life Satisfaction" = c(
    "health", "lifesat", "headache", "stomachache", "backache",
    "feellow", "irritable", "nervous", "sleepdificulty", "dizzy"
  ),
  "Problematic Social Media Use (PMSU)" = c(
    "emcsocmed1", "emcsocmed2", "emcsocmed3",
    "emcsocmed4", "emcsocmed5", "emcsocmed6",
    "emcsocmed7", "emcsocmed8", "emcsocmed9"
  ),
  "Bullying" = c(
    "beenbullied", "cbeenbullied"
  ),
  "Communicates Online More Than Real Life" = c(
    "emconlfreq1", "emconlfreq2", "emconlfreq3", "emconlfreq4",
    "emconlpref1", "emconlpref2", "emconlpref3"
  ),
  "Family Support" = c(
    "famhelp", "famsup", "famtalk", "famdec"
  ),
  "Friends Support" = c(
    "friendhelp", "friendcounton", "friendshare", "friendtalk"
  ),
  "School Support" = c(
    "likeschool", "schoolpressure",
    "studtogether", "studhelpful", "studaccept",
    "teacheraccept", "teachercare", "teachertrust"
  ),
  "Communication with Parents" = c(
    "talkfather", "talkmother"
  ),
  "Good/Bad Habits" = c(
    "physact60", "timeexe", "alc30d_2"
  )
)

# create dat
cols_to_keep <- unlist(var_groups, use.names = FALSE)
dat <- hbsc2018 %>%
  select(all_of(cols_to_keep))

# prettt print vars
 for (group_name in names(var_groups)) {
   cat("##", group_name, "\n")
 
   cat(
     paste0(
       "* `", var_groups[[group_name]], "`",
       collapse = "\n"
     ),
     "\n\n"
   )
 }

```

The number of missing values is relevant and need to address later:

```{r nas, echo=FALSE}
na_cnts <- data.frame(
  variable = names(dat),
  na_count = sapply(dat, function(x) sum(is.na(x)))
)

row.names(na_cnts) <- NULL

na_cnts <- na_cnts %>%
  #mutate(na_pct = round(100 * na_count / nrow(dat),0)) %>%
  mutate(na_pct = na_count / nrow(dat)) %>% 
  mutate(na_pct = percent(na_pct, accuracy = 1)) %>%
  arrange(desc(na_count))

print(na_cnts)
```

Now I explore all variables by group and feature engineer as appropriate.

# [1] "Individual Information"                 

11 vars here. So do 1x1.

```{r ii, echo=FALSE}

vars_indivinfo <- var_groups[["Individual Information"]]
vars_indivinfo

```

seqno_int is just an identifier.

```{r, include=FALSE}

length(table(dat$seqno_int)) == nrow(dat)

```

countryno is recoded to country_name and this is the distribution.

```{r country, echo=FALSE}

recode_map <- c(
  "8000"   = "Albania",
  "31000"  = "Azerbaijan",
  "40000"  = "Austria",
  "51000"  = "Armenia",
  "56001"  = "Belgium (Flemish)",
  "56002"  = "Belgium (French)",
  "100000" = "Bulgaria",
  "124000" = "Canada",
  "191000" = "Croatia",
  "203000" = "Czech Republic",
  "208000" = "Denmark",
  "233000" = "Estonia",
  "246000" = "Finland",
  "250000" = "France",
  "268000" = "Georgia",
  "276000" = "Germany",
  "300000" = "Greece",
  "304000" = "Greenland",
  "348000" = "Hungary",
  "352000" = "Iceland",
  "372000" = "Ireland",
  "376000" = "Israel",
  "380000" = "Italy",
  "398000" = "Kazakhstan",
  "428000" = "Latvia",
  "440000" = "Lithuania",
  "442000" = "Luxembourg",
  "470000" = "Malta",
  "498000" = "Republic of Moldova",
  "528000" = "Netherlands",
  "578000" = "Norway",
  "616000" = "Poland",
  "620000" = "Portugal",
  "642000" = "Romania",
  "643000" = "Russia",
  "688000" = "Serbia",
  "703000" = "Slovakia",
  "705000" = "Slovenia",
  "724000" = "Spain",
  "752000" = "Sweden",
  "756000" = "Switzerland",
  "792000" = "Turkey",
  "804000" = "Ukraine",
  "807000" = "Macedonia",
  "826001" = "England",
  "826002" = "Scotland",
  "826003" = "Wales",
  "826004" = "Northern Ireland",
  "840000" = "USA"
)

dat$country_name <- recode_map[as.character(dat$countryno)]

tbl_country <- dat %>%
  group_by(country_name) %>%
  summarise(n = n(),
            .groups = "drop") %>%
  #mutate(pct = round(100 * n / sum(n),1)) %>%
  mutate(pct = n / sum(n)) %>%  
  mutate(pct = percent(pct, accuracy = 1)) %>% 
  arrange(country_name) 

kable(tbl_country,
      caption = "#/% of observations by country")

dat <- dat %>% 
  select(-countryno)

```

bodyheight is in cm and range from 120 to 200 so will use. Impute NAs?

```{r, echo=FALSE}

summary(dat$bodyheight)

ggplot(dat, aes(x = bodyheight)) +
  geom_histogram() +
  #facet_grid(sex ~ .) +
  theme_minimal()

```

bodyweight is in kilo and range from 20 to 150 so will use maybe. Impute NAs?

```{r, echo=FALSE}

summary(dat$bodyweight)

ggplot(dat, aes(x = bodyweight)) +
  geom_histogram() +
  #facet_grid(sex ~ .) +
  theme_minimal()

```

MBMI is body mass index from [Range= 11.02-44.9] and 0 meaning outside overall range so make those NA.

```{r, echo=FALSE}

summary(dat$MBMI)

dat <- dat %>% 
  mutate(MBMI = ifelse(MBMI==0, NA, MBMI))

ggplot(dat, aes(x = MBMI)) +
  geom_histogram() +
  #facet_grid(sex ~ .) +
  theme_minimal()

```

IOTF4 is a BMI group from 1(thin) to 4(obesse) so leave as is.

```{r iot, echo=FALSE}

table(dat$IOTF4, useNA = "always")

```

agecat has 3 values and will leave as is: 1 ~ 11, 2 ~ 13, 3 ~ 15.

```{r age, echo=FALSE}

table(dat$agecat, useNA = "always")

```

sex is binary with 1 (boy), 2 (girl) so will leave as is.

```{r sex, echo=FALSE}

table(dat$sex, useNA = "always")

```

IRRELFAS_LMH is relative family affluence with 1 Low20, 2 Mid60, 3 High20. Leave as is.

IRFAS is discrete [0,13] and shows affluence.

```{r fas, echo=FALSE}

#table(dat$IRRELFAS_LMH, useNA = "always")

#table(dat$IRFAS)

table(dat$IRFAS,
      dat$IRRELFAS_LMH,
      useNA = "always")

```

# [2] "Health Issues and Life Satisfaction"   
10 vars here:

```{r hea, echo=FALSE}

var_groups[[2]]

```
health is self rating of health from 1(Excellent) to 4(poor). Won't use for now

```{r}

table(dat$health, useNA = "always")

```

lifesat asks for life satisfaction as ladder where 0(botton) to 10(best). I'll dichotomize and keep num for shap.

```{r lifesat, echo=FALSE}

cat("lifesat distribution: ")
table(dat$lifesat, useNA = "always")

cat("lifesat %: ")
round(table(dat$lifesat, useNA = "always") / nrow(dat),3)

dat <- dat %>%
  mutate(
    lifesat_low = case_when(
      between(lifesat, 0, 5) ~ 1,
      between(lifesat, 6, 10) ~ 0,
      TRUE ~ NA))

cat("lifesat_low as y/n: ")
table(dat$lifesat,
      dat$lifesat_low,
      useNA = "always"
      )

```

headache, stomachache, backache, and dizzy are physical and will ignore for now. But many papers report on psychosomatic issues so I might come back.

feellow, irritable, nervous, sleepdificulty and mental health issues. They answer a question In the last 6 months: how often have you had the following….? And the answers goes from 1 (About every day) to 5 (Rarely or never). Their distribution is:

```{r vm1, echo=FALSE}

vars_mental <- c("feellow", "irritable", "nervous", "sleepdificulty")

# tabulate the tidy way with purrr
tables_list <- map(1:4, ~ table(dat[[vars_mental[.x]]]))
names(tables_list) <- vars_mental[1:4]
print(tables_list)

```

Their correlation is modest.

```{r mhcor, echo=FALSE}

mh_cor <- cor(dat[vars_mental],
              use = "pairwise.complete.obs")
mh_cor <- round(mh_cor, 2)
print(mh_cor)

```

These variables are then recoded to a binary where it flags if the complaint
occurs more than once per week (values 1/2).

```{r mrrec, echo=FALSE}

dat <- dat %>%
  mutate(
#    headache_frequent = ifelse(headache <= 2, 1, 0),
#    stomachache_frequent = ifelse(stomachache <= 2, 1, 0),
#    backache_frequent = ifelse(backache <= 2, 1, 0),
#    dizzy_frequent = ifelse(dizzy <= 2, 1, 0),
    feellow_is = ifelse(feellow <= 2, 1, 0),
    irritable_is = ifelse(irritable <= 2, 1, 0),
    nervous_is = ifelse(nervous <= 2, 1, 0),
    sleepdificulty_is = ifelse(sleepdificulty <= 2, 1, 0)
  )

```

These frequency 1/0 vars are summed up and if sum is 2+ then there's an issue. And counts look like this:

```{r mh_rec, echo=FALSE}

dat <- dat %>%
  # add the frequency of complaints
  mutate(mental_sum = feellow_is +
                      irritable_is +
                      nervous_is +
                      sleepdificulty_is
  ) %>%
  # derive index based on sum
  mutate(mental_issue = ifelse(mental_sum >=2, 1, 0)#,
      ) 

table(dat$mental_sum,
      dat$mental_issue,
      useNA = "always")

```

# [3] "Problematic Social Media Use (PMSU)"  
9 binary variables are recoded to 0/1.

```{r pmsu_recode, echo=FALSE}

vars_pmsu <- var_groups[["Problematic Social Media Use (PMSU)"]]
vars_pmsu_r <- paste0(vars_pmsu, "_r")
  
# recode to 0/1 from 1/2
recode_emcsocmed <- function(old_col){
  new_col <- case_when(
    old_col == 1 ~ 0,
    old_col == 2 ~ 1,
    TRUE ~ NA #99?
  )
  return(new_col)
}

dat <- dat %>%
  mutate(emcsocmed1_r = recode_emcsocmed(emcsocmed1),
         emcsocmed2_r = recode_emcsocmed(emcsocmed2),
         emcsocmed3_r = recode_emcsocmed(emcsocmed3),
         emcsocmed4_r = recode_emcsocmed(emcsocmed4),
         emcsocmed5_r = recode_emcsocmed(emcsocmed5),
         emcsocmed6_r = recode_emcsocmed(emcsocmed6),
         emcsocmed7_r = recode_emcsocmed(emcsocmed7),
         emcsocmed8_r = recode_emcsocmed(emcsocmed8),
         emcsocmed9_r = recode_emcsocmed(emcsocmed9)
         ) 

# and look like pprrint:
questions <- data.frame(
  item = c(
    "emcsocmed1","emcsocmed2","emcsocmed3","emcsocmed4",
    "emcsocmed5","emcsocmed6","emcsocmed7","emcsocmed8","emcsocmed9"
  ),
  topic = c(
    "Social media: I can't think of anything else",
    "Social media: Dissatisfied cause want to spend more time",
    "Social media: Felt bad cause want to play",
    "Social media: Failed to spend less time",
    "Social media: Neglected other activities",
    "Social media: Had arguments because of use",
    "Social media: Lied about amount of time spent",
    "Social media: Use it to escape from negative feelings",
    "Social media: Conflict with family because of use"
  )
)

kable(questions) #, caption = "Social Media Use Items")

```

And they are correlated.

```{r corr2,echo=FALSE}

tetra_mat <- tetrachoric(dat[vars_pmsu_r])$rho
tetra_mat <- round(tetra_mat, 2)

corrplot::corrplot(
  tetra_mat,
  method = "color",
  type = "upper",
  tl.col = "black",
  addCoef.col = "black"
)

```

These 9 vars are sumed and then binned to H/M/L where H(6-9), M(2-5) and L(0-1). Resulting in:

```{r pmsu_recode2, echo=FALSE}

#sum and bin
dat <- dat %>%
  # should go from [0,9]
  mutate(pmsu_sum = rowSums(dat[,vars_pmsu_r])) %>%
  # 3 groups
  mutate(
    pmsu_lmh = case_when(
      between(pmsu_sum,6,9) ~ 3, #"High",
      between(pmsu_sum,2,5) ~ 2, #"Med",
      between(pmsu_sum,0,1) ~ 1, #"Low",
      TRUE ~ NA
    )
  )

#cat("## Table for emcsocmed_sum x pmsu_lmh")
kable(table(dat$pmsu_sum,
      dat$pmsu_lmh,
      useNA = "always")
)

# remove original vars
dat <-  dat %>%
  select(-all_of(vars_pmsu))

```


# [4] "Bullying"                               

There are 2 variables that measure if kid has been bullied. a) beenbullied: Been bullied past months
and b) cbeenbullied: Been cyber bullied. Answers range from 1 (Haven't) to 5 (Several times a week).  

I will derive 2 to keep: cbeenbullied as y/n and bulliedsum as sum of both.

```{r bully, echo=FALSE}

dat <- dat %>%
  mutate(cbullied = 
           case_when(cbeenbullied == 1 ~ 0,
                     between(cbeenbullied, 2, 5) ~ 1,
                     TRUE ~ NA),
         bulliedsum = beenbullied + cbeenbullied
           )

cat("cbullied recoding:")
table(dat$cbeenbullied,
      dat$cbullied,
      useNA = "always")

bullydf <- data.frame(table(dat$bulliedsum, useNA = "always")) %>%
  rename(score = Var1,
         n = Freq) %>%
  mutate(pct = n / sum(n),
         pct = percent(pct, accuracy = 1))

cat("bullying score: (low is better)")
kable(bullydf)

```

# [5] "Communicates Online More Than Real Life"  

7 vars here in 2 themes - a) continuos online communications and b) preference to communicate online.

For a) continuos online communications, measures how how often young person communicate online with 
4 types of people in their life. Answers range from 1 (NA or don't know) to 6 (almost all of the time). 

| Variable      | Description                         |
|---------------|-------------------------------------|
| emconlfreq1   | Online contact with close friends    |
| emconlfreq2   | Online contact with larger friend group |
| emconlfreq3   | Online contact with online friends   |
| emconlfreq4   | Online contact with other            |

And they are somewhat correlated but not much.

```{r corr_emconlfreqX,echo=FALSE}

vars_onlfe <- var_groups[["Communicates Online More Than Real Life"]]
vars_onlfe

corr_emconlfreqX <- cor(dat[vars_onlfe[1:4]],
                        use = "pairwise.complete.obs")

corr_emconlfreqX <- round(corr_emconlfreqX, 2)

corrplot::corrplot(
  corr_emconlfreqX,
  method = "color",
  type = "upper",
  tl.col = "black",
  addCoef.col = "black"
)

```

I combine them into two options. a) one dichotomized and adopt the HBSC definition of yes if 
any of the variables is 6; and b) a sum for our shap tests with values in [4,24].  

```{r livingonline, echo=FALSE}

dat <- dat %>%
  mutate(
    # 1 if any of the 4 is 6
    onlfreq_is = case_when(
      (emconlfreq1 == 6 | emconlfreq2 == 6 | emconlfreq3 == 6 | emconlfreq4 == 6) ~ 1,
      (!is.na(emconlfreq1) | !is.na(emconlfreq2) | !is.na(emconlfreq3) | !is.na(emconlfreq4)) ~ 0,
      TRUE ~ NA #still need to do something with 99
    ),
    # sums the 4
    onlfreq_sum = emconlfreq1 + emconlfreq2 + emconlfreq3 + emconlfreq4
  )
    
cat("Online frequence is (onlfreq_is): ")
table(dat$onlfreq_is, useNA = "always")

summary(dat$onlfreq_sum)

ggplot(dat, aes(x=onlfreq_sum)) +
  geom_bar() +
  theme_minimal()

```

For b) preference to communicate online, 3 questions ask about preference of sharing important stuff online vs real-life. The three variables are:

| Variable      | Description                     |
|---------------|---------------------------------|
| emconlpref1   | Secrets, more easily online     |
| emconlpref2   | Feelings, more easily online    |
| emconlpref3   | Concerns, more easily online    |

The answers range from 1 (Strong disagree) to 5 (Strong agree).

And they are correlated. Remove 99s????

```{r corr_emconlprefX,echo=FALSE}

#vars_emconlpref <- var_groups[["Communicates Online More Than Real Life"]]
#vars_onlfe

corr_emconlprefX <- cor(dat[vars_onlfe[5:7]],
                        use = "pairwise.complete.obs")

corr_emconlprefX <- round(corr_emconlprefX, 2)

corrplot::corrplot(
  corr_emconlprefX,
  method = "color",
  type = "upper",
  tl.col = "black",
  addCoef.col = "black"
)

```

Haven't found a paper that uses them so I will replicated method of friends support below. Summing them results in [3,15] and then grouped as H/M/L as L 3-6, M 7-10, H 11-15. Resulting in:

```{r emconlprefX, echo=FALSE}

dat <- dat %>%
  mutate(# sum should give 3-15
        onlpref_sum = emconlpref1 + emconlpref2 + emconlpref3,
        onlpref_sum = ifelse(onlpref_sum > 15, NA, onlpref_sum), #for now, 99 issue
        # create hml
        onlpref = case_when(
          between(onlpref_sum, 3, 6) ~ 1,
          between(onlpref_sum, 7, 10) ~ 2,
          between(onlpref_sum, 11, 15) ~ 3,
          TRUE ~ NA
        )
  )
    
cat("onlpref: ")
table(dat$onlpref, useNA = "always")

summary(dat$onlpref_sum)

ggplot(dat, aes(x=onlpref_sum)) +
  geom_bar() +
  theme_minimal()

```

# [6] "Family Support" 
4 questions going from 1 Very strongly disagree to 7 Very strongly agree.

```{r famprnt, echo=FALSE}

vars_famsup <- var_groups[["Family Support"]]
vars_famsup

family_items <- data.frame(
  item = c(
    "famhelp",
    "famsup",
    "famtalk",
    "famclose"
  ),
  topic = c(
    "Family tries to help",
    "Get emotional help from family",
    "Talk about problems with family",
    "Feel close to family"
  )
)

kable(family_items) #, caption = "Family Support Items")
```

And they are correlated.

```{r famcorr2,echo=FALSE}

#tetra_mat <- tetrachoric(dat[vars_famsup])$rho
cor_famsup <- cor(dat[vars_famsup],
    use = "pairwise.complete.obs")

#tetra_mat <- round(tetra_mat, 2)

corrplot::corrplot(
  cor_famsup,
  method = "color",
  type = "upper",
  tl.col = "black",
  addCoef.col = "black"
)

```

Summing them results in [4,28] and then grouped as H/M/L as L 4-11, M 12-19, H 20-28. Resulting in:

```{r fam_rec, echo=FALSE}

dat <- dat %>%
   mutate(famsup_sum = rowSums(dat[,vars_famsup])
          ) %>%
   mutate(famsup_hml = case_when(
            famsup_sum >= 4 & famsup_sum <= 11 ~ 1, #"Low",
            famsup_sum >= 12 & famsup_sum <= 19 ~ 2, #"Med",
            famsup_sum >= 20 & famsup_sum <= 28 ~ 3, #"High",
            TRUE ~ NA))

tbl_famsup <- table(dat$famsup_sum,
                    dat$famsup_hml,
                    useNA = "always"
                    )

kable(tbl_famsup)

```


# [7] "Friends Support"   
4 questions going from 1 Very strongly disagree to 7 Very strongly agree.

```{r friprnt, echo=FALSE}

vars_frisup <- var_groups[["Friends Support"]]
vars_frisup

friend_items <- data.frame(
  item = c(
    "friendhelp",
    "friendcounton",
    "friendshare",
    "friendtalk"
  ),
  topic = c(
    "My friends really try to help me",
    "I can count on my friends when things go wrong",
    "I have friends with whom I can share my joys and sorrows",
    "I can talk about my problems with my friends"
  )
)

kable(friend_items)

```

And they are correlated.

```{r frisup_r,echo=FALSE}

#tetra_mat <- tetrachoric(dat[vars_famsup])$rho
cor_frisup <- cor(dat[vars_frisup],
    use = "pairwise.complete.obs")

#tetra_mat <- round(tetra_mat, 2)

corrplot::corrplot(
  cor_frisup,
  method = "color",
  type = "upper",
  tl.col = "black",
  addCoef.col = "black"
)

```

Summing them results in [4,28] and then grouped as H/M/L as L 4-11, M 12-19, H 20-28. Resulting in:

```{r fri_rec, echo=FALSE}

dat <- dat %>%
   mutate(frisup_sum = rowSums(dat[,vars_frisup])
          ) %>%
   mutate(frisup_hml = case_when(
            frisup_sum >= 4 & frisup_sum <= 11 ~ 1, #"Low",
            frisup_sum >= 12 & frisup_sum <= 19 ~ 2, #"Med",
            frisup_sum >= 20 & frisup_sum <= 28 ~ 3, #"High",
            TRUE ~ NA))

tbl_frisup <- table(dat$frisup_sum,
                    dat$frisup_hml,
                    useNA = "always"
                    )

kable(tbl_frisup)

```

# [8] "School Support"                         
8 variables related to school.

```{r sc1, echo=FALSE}
var_groups[8]

```
likeschool ask if they like school and goes from 1 (lots) to 4 (no). I won't use for now.

schoolpressure ask if they feel pressure and goes from 1(no) to 4(a lot). I won't use for now.

Then there are 3 questions about relationship with students and 3 with relationship with teachers. The answers go from 1 (strong agree) to 5 (strong disagree).

```{r skool1, echo=FALSE}

school_items <- data.frame(
  category = c("teacher", "teacher", "teacher",
               "classmates", "classmates", "classmates"),
  item = c("teacheraccept",
           "teachercare",
           "teachertrust",
           "studtogether",
           "studhelpful",
           "studaccept"),
  topic = c("Teacher accepts me",
            "Teacher cares about me",
            "Feel trust in teacher",
            "Students enjoy being together",
            "Students kind and helpful",
            "Students accept me")
)

kable(school_items)


```

Scales are reversed so 1 is the negative and 5 is the positive.

```{r skurecode, echo=FALSE}

# recodes from 1-5 to 0-4 
recode_school_vars <- function(old_values){
  new_values <- case_when(
    old_values == 5 ~ 1,
    old_values == 4 ~ 2,
    old_values == 3 ~ 3,
    old_values == 2 ~ 4,
    old_values == 1 ~ 5,
    TRUE ~ NA
  )
  return(new_values)
}

dat <- dat %>%
  # recode so they can be added
  mutate(teacheraccept_r = recode_school_vars(teacheraccept),
         teachercare_r = recode_school_vars(teachercare),
         teachertrust_r = recode_school_vars(teachertrust),
         studtogether_r = recode_school_vars(studtogether),
         studhelpful_r = recode_school_vars(studhelpful),
         studaccept_r = recode_school_vars(studaccept)
         ) 

```

And there is moderate correlation between groups but lowers across stud/teach.

```{r skulr, echo=FALSE}

vars_school <- c("teacheraccept", "teachercare","teachertrust",
                 "studtogether", "studhelpful", "studaccept")
vars_school_r <- paste0(vars_school, "_r")

school_r <- cor(dat[vars_school_r],
              use = "pairwise.complete.obs")
print(round(school_r,2))

```

A few features are derived. For stud/teach separate a binary following HBSC and a sum for shap test.

```{r skulfe, echo=FALSE}

dat <- dat %>%
  # derive avg by teacher/student
  mutate(teacher_support_avg = (teacheraccept_r 
                               + teachercare_r
                               + teachertrust_r) / 3,
         student_support_avg = (studtogether_r
                                + studhelpful_r
                                + studaccept_r) / 3
         ) %>%
  # derive support yn based on hbsc
  mutate(
    teachersup = case_when(
      teacher_support_avg >= 4 ~ 1,
      teacher_support_avg < 4 ~ 0,
      TRUE ~ NA),
      studsup = case_when(
        student_support_avg >= 4 ~ 1,
        student_support_avg < 4 ~ 0,
        TRUE ~ NA)
  ) %>%
  # derive support sum for shap
  mutate(teachersup_sum = teacheraccept_r 
                            + teachercare_r
                            + teachertrust_r,
         studsup_sum = studtogether_r
                        + studhelpful_r
                        + studaccept_r
         )

cat ("teachersup: ")
table(dat$teachersup_sum,
      dat$teachersup,
      useNA = "always")

cat ("studentsup: ")
table(dat$studsup_sum,
      dat$studsup,
      useNA = "always")

# clean a bit
dat <- dat %>%
  select(-c(teacher_support_avg, student_support_avg),
         all_of(vars_school))

```

# [9] "Communication with Parents"

Two vars with answers from 1 (Very easy) to 5 (Don't have or see). With this counts and proportions in data.  

```{r commp, echo=FALSE}

vars_commparents <- var_groups[["Communication with Parents"]]
vars_commparents

table(dat[vars_commparents], useNA = "always")
round(table(dat[vars_commparents], useNA = "always") / nrow(dat),3)

```

With some correlation between them (I expected more).

```{r cpcorr, echo=FALSE}

corrmat <- cor(dat[vars_commparents], 
               use = "pairwise.complete.obs")

```

So I will dichotomize like HBSC and keep sum of both as well for SHAP. These are dichotomized for each parent as Yes if [1,2] and No if [3-5].  

```{r, include=FALSE}

dat <- dat %>%
  mutate(talkf = 
           case_when(talkfather %in% c(1,2) ~ 1,
                     talkfather %in% c(3,4,5) ~ 0,
                     TRUE ~ NA),
         talkm = 
           case_when(talkmother %in% c(1,2) ~ 1,
                     talkmother %in% c(3,4,5) ~ 0,
                     TRUE ~ NA),
         talkscore = talkfather + talkmother
         )

```

Where new vars look like this for dichotom values:

```{r prcom, echo=FALSE}

cat("Father:")
table(dat$talkfather,
      dat$talkf,
      useNA = "always")

cat("Mother:")
table(dat$talkfather,
      dat$talkf,
      useNA = "always")

```

And the talk score looks like this. Will keep for now and orig vars as well.

```{r prsum, echo=FALSE}

#summary(dat$talkscore)

ggplot(dat, aes(x = factor(talkscore))) +
  geom_bar() +
  #scale_x_discrete() +
  theme_minimal()

```

# [10] "Good/Bad Habits"

physact60 is just past week so ignore.

timeexe ask how often exercise in free time going from 1 (every day) to 7 never. HBSC shows proportions of those that do 3+ per week. But that is 2022 data, in our case we do 2-3 times per week or more. And I keep timexe as is for shap test.

```{r timeexe_rec, echo=FALSE}

dat <- dat %>%
  mutate(timeexe_r = case_when(
    between(timeexe, 1, 3) ~ 1,
    between(timeexe, 4, 7) ~ 0,
    TRUE ~ NA
  ))

print(
table(dat$timeexe,
      dat$timeexe_r,
      useNA = "ifany")
)

```

alc30d_2 ask about alcohol use in past 30 days and will ignore for now

# Clean Dataset for Modeling

```{r dclean, include=FALSE}

glimpse(dat)

# Create the structured list of variables
vars_list_mdls <- list(
  personal = c(
    "seqno_int", "country_name", "bodyheight", "bodyweight", 
    "MBMI", "IOTF4", "agecat", "sex", "IRRELFAS_LMH", "IRFAS"
  ),
  health = c(
    "lifesat_low", "mental_issue", "mental_sum"
  ),
  pmsu = c(
    "pmsu_lmh", "pmsu_sum"
  ),
  bullying = c(
    "cbullied", "bulliedsum"
  ),
  online_comms = c(
    "onlfreq_is", "onlfreq_sum", "onlpref", "onlpref_sum"
  ),
  afmsup = c(
    "famsup_hml", "famsup_sum"
  ),
  friend_support = c(
    "frisup_hml", "frisup_sum"
  ),
  school_support = c(
    "teachersup", "teachersup_sum", "studsup", "studsup_sum"
  ),
  parent_comms = c(
    "talkf", "talkm", "talkscore"
  ),
  habits = c(
    "timeexe_r", "timeexe"
  )
)

vars_mdls <- unlist(vars_list_mdls, use.names = FALSE)
d <- dat %>% select(all_of(vars_mdls))

# How to select #1 personal:
#personal_vars <- variables_list$personal
#print(personal_vars)

```

# Missing Values
Need to plan for this. For now, just keeping complete obs, but note some countries are very affected. 

```{r na, echo=FALSE}

d <- d %>% 
  mutate(any_na = if_any(everything(), is.na))

d %>%
  group_by(country_name, any_na) %>%
  tally() %>%
  pivot_wider(names_from = any_na,
              names_prefix = "na_",
              values_from = n,
              values_fill = 0) %>%
  mutate(n = na_FALSE + na_TRUE) %>%
  mutate(pct_complete = round(na_FALSE / n,3)) %>%
  select(n, na_FALSE, pct_complete) %>%
  arrange(desc(pct_complete))-> tbl_nas_country

kable(tbl_nas_country)

d <- d[complete.cases(d),] %>%
  select(-any_na)

#summary(dc)

```

# Data Type Coercions

```{r}

# # vars to coerce to fact (i looked 1x1)
# country_name
# IOTF4 ordered 1-4
# agecat ordered 1-3
# sex
# IRRELFAS_LMH ordered 1-3
# lifesat_low ordered 0-1
# mental_issue ordered 0-1
# pmsu_lmh order 1-3
# cbullied order 0-1
# onlfreq_is order 0-1
# onlpref order 1-3
# famsup_hml order 1-3
# frisup_hml order 1-3
# teachersup order 0-1
# studsup order 0-1
# talkf order 0-1
# talkm order 0-1
# timeexe_r order 0-1

d <- d %>%
  mutate(
    # --- Unordered (Nominal) Factors ---
    country_name = as.factor(country_name),
    sex          = as.factor(sex),
    # --- Ordered Factors (1 to 4) ---
    IOTF4        = factor(IOTF4, levels = c("1", "2", "3", "4"), ordered = TRUE),
    # --- Ordered Factors (1 to 3) ---
    agecat       = factor(agecat,       levels = c("1", "2", "3"), ordered = TRUE),
    IRRELFAS_LMH = factor(IRRELFAS_LMH, levels = c("1", "2", "3"), ordered = TRUE),
    pmsu_lmh     = factor(pmsu_lmh,     levels = c("1", "2", "3"), ordered = TRUE),
    onlpref      = factor(onlpref,      levels = c("1", "2", "3"), ordered = TRUE),
    famsup_hml   = factor(famsup_hml,   levels = c("1", "2", "3"), ordered = TRUE),
    frisup_hml   = factor(frisup_hml,   levels = c("1", "2", "3"), ordered = TRUE),
    # --- Ordered Factors (0 to 1 Binary / Dummy Scales) ---
    lifesat_low  = factor(lifesat_low,  levels = c("0", "1"), ordered = TRUE),
    mental_issue = factor(mental_issue, levels = c("0", "1"), ordered = TRUE),
    cbullied     = factor(cbullied,     levels = c("0", "1"), ordered = TRUE),
    onlfreq_is   = factor(onlfreq_is,   levels = c("0", "1"), ordered = TRUE),
    teachersup   = factor(teachersup,   levels = c("0", "1"), ordered = TRUE),
    studsup      = factor(studsup,      levels = c("0", "1"), ordered = TRUE),
    talkf        = factor(talkf,        levels = c("0", "1"), ordered = TRUE),
    talkm        = factor(talkm,        levels = c("0", "1"), ordered = TRUE),
    timeexe_r    = factor(timeexe_r,    levels = c("0", "1"), ordered = TRUE)
  )

```

# Tests for Potential of Prediction

```{r predpowertests}

#vars_for cochran (ordered)
#vars_for_chi
#vars_for_tt 

# --- 1. SETUP YOUR TARGET VARIABLE NAME ---
# Change "depression" to your exact binary response column name
# IMPORTANT: This target variable must have exactly two levels (e.g., "No" and "Yes")
#target_var <- "depression"
target_var <- "mental_issue"
# Ensure target is a factor and identify the "success" level (the 2nd level)
#d[[target_var]] <- as.factor(d[[target_var]])
target_levels <- levels(d[[target_var]])
success_level <- target_levels[2] 

# --- 2. THE AUTOMATED LOOP ---
for (col_name in names(d)) {
  
  # Skip the target variable itself
  if (col_name %in% c(target_var, "seqno_int")) next
  
  cat("\n==================================================\n")
  cat("Testing Relationship:", target_var, "vs", col_name, "\n")
  cat("==================================================\n")
  
  # --- 1. IF ORDINAL -> RUN COCHRAN-ARMITAGE TREND TEST ---
  if (is.ordered(d[[col_name]])) {
    cat("Predictor Type: ORDINAL -> Running Cochran-Armitage Trend Test...\n\n")
    
    # Build table: rows = target, columns = ordinal levels
    contingency_table <- table(d[[target_var]], d[[col_name]])
    print(contingency_table)
    cat("\n")
    
    # Extract successes for the 2nd level (e.g., "Yes") and column totals
    successes <- contingency_table[success_level, ]
    totals    <- colSums(contingency_table)
    
    test_result <- tryCatch({
      prop.trend.test(successes, totals)
    }, error = function(e) { return(NULL) })
    
    if (!is.null(test_result)) {
      print(test_result)
    } else {
      cat("Error: Could not run Trend Test. Verify your variable setup.\n")
    }
    
  # --- 2. IF NUMERIC -> RUN T-TEST ---
  } else if (is.numeric(d[[col_name]])) {
    cat("Predictor Type: NUMERIC -> Running Independent t-test...\n\n")
    
    test_result <- tryCatch({
      t.test(d[[col_name]] ~ d[[target_var]])
    }, error = function(e) { return(NULL) })
    
    if (!is.null(test_result)) {
      print(test_result)
    } else {
      cat("Error: Could not run t-test. Check for missing values.\n")
    }
    
  # --- 3. IF NOMINAL / BINARY -> RUN CHI-SQUARE ---
  } else if (is.factor(d[[col_name]]) || is.character(d[[col_name]])) {
    cat("Predictor Type: NOMINAL/BINARY -> Running Chi-Square Test...\n\n")
    
    contingency_table <- table(d[[target_var]], d[[col_name]])
    print(contingency_table)
    cat("\n")
    
    test_result <- tryCatch({
      chisq.test(contingency_table)
    }, error = function(e) { return(NULL) })
    
    if (!is.null(test_result)) {
      print(test_result)
    } else {
      cat("Error: Could not run Chi-Square test.\n")
    }
  }
}


# # chi for nominal
# cont_tbl <- table(d[["mental_issue"]], d[["pmsu_lmh"]])
# print(cont_tbl)
# chi <- chisq.test(cont_tbl)
# print(chi)
# print(chi$p.value)
# print(chi$statistic)
# 
# # cochran for ordered
# cont_tbl <- table(d[["mental_issue"]], d[["pmsu_lmh"]])
# cont_tbl
# successes <- cont_tbl["1", ]
# successes
# totals <- colSums(cont_tbl)
# coch <- prop.trend.test(successes, totals)
# coch
# 
# # t.test for num
# tt <- t.test(d[["pmsu_sum"]] ~ d[["mental_issue"]])
# tt    


    



```


```{r}
# --- 1. SETUP YOUR TARGET VARIABLE NAME ---
# Target variable is set to "mental_issue" and dataset is named "d"
target_var <- "mental_issue"

# Ensure target is a factor and identify the "success" level (the 2nd level)
d[[target_var]] <- as.factor(d[[target_var]])
target_levels <- levels(d[[target_var]])
success_level <- target_levels[2] # Grabs group 1 / "Yes" status automatically

# --- 2. INITIALIZE AN EMPTY DATAFRAME FOR THE SUMMARY TABLE ---
summary_table <- data.frame(
  Predictor_Variable = character(),
  Variable_Type      = character(),
  Statistical_Test   = character(),
  Test_Statistic     = character(),
  p_value            = character(),
  stringsAsFactors   = FALSE
)

# --- 3. THE AUTOMATED LOOP ---
for (col_name in names(d)) {
  
  # Skip the target variable itself
  if (col_name == target_var) next
  
  # Create baseline variables for storing this row's data
  var_type <- "Unknown"
  test_name <- "None"
  stat_val <- "N/A"
  p_val_str <- "N/A"
  
  # --- 1. IF ORDINAL -> RUN COCHRAN-ARMITAGE TREND TEST ---
  if (is.ordered(d[[col_name]])) {
    var_type <- "Ordinal"
    test_name <- "Cochran-Armitage Trend Test"
    
    contingency_table <- table(d[[target_var]], d[[col_name]])
    successes <- contingency_table[success_level, ]
    totals    <- colSums(contingency_table)
    
    test_result <- tryCatch({ prop.trend.test(successes, totals) }, error = function(e) { return(NULL) })
    
    if (!is.null(test_result)) {
      stat_val <- paste0("X-squared = ", round(test_result$statistic, 2))
      p_val_str <- if(test_result$p.value < 0.001) "< 0.001" else format(round(test_result$p.value, 4), scientific = FALSE)
    }
    
  # --- 2. IF NUMERIC -> RUN T-TEST ---
  } else if (is.numeric(d[[col_name]])) {
    var_type <- "Numeric"
    test_name <- "Welch Two Sample t-test"
    
    test_result <- tryCatch({ t.test(d[[col_name]] ~ d[[target_var]]) }, error = function(e) { return(NULL) })
    
    if (!is.null(test_result)) {
      stat_val <- paste0("t(", round(test_result$parameter, 1), ") = ", round(test_result$statistic, 2))
      p_val_str <- if(test_result$p.value < 0.001) "< 0.001" else format(round(test_result$p.value, 4), scientific = FALSE)
    }
    
  # --- 3. IF NOMINAL / BINARY -> RUN CHI-SQUARE ---
  } else if (is.factor(d[[col_name]]) || is.character(d[[col_name]])) {
    var_type <- if(length(unique(d[[col_name]])) == 2) "Binary" else "Nominal"
    test_name <- "Pearson's Chi-Square Test"
    
    contingency_table <- table(d[[target_var]], d[[col_name]])
    test_result <- tryCatch({ chisq.test(contingency_table) }, error = function(e) { return(NULL) })
    
    if (!is.null(test_result)) {
      stat_val <- paste0("X-squared(", test_result$parameter, ") = ", round(test_result$statistic, 2))
      p_val_str <- if(test_result$p.value < 0.001) "< 0.001" else format(round(test_result$p.value, 4), scientific = FALSE)
    }
  }
  
  # --- 4. APPEND THE RESULT ROW TO OUR TABLE ---
  new_row <- data.frame(
    Predictor_Variable = col_name,
    Variable_Type      = var_type,
    Statistical_Test   = test_name,
    Test_Statistic     = stat_val,
    p_value            = p_val_str,
    stringsAsFactors   = FALSE
  )
  summary_table <- rbind(summary_table, new_row)
}

# --- 4. VIEW THE FINAL SUMMARY TABLE ---
print(summary_table)

```

```{r my1sttidyLR}

# # create modeling dat
# dat <- dat %>%
#   mutate(y = ifelse(lifesat >5,1,0),
#          y = factor(y))

d_mdl <- d %>% 
  select(country_name, MBMI, agecat, sex, IRRELFAS_LMH, mental_issue, pmsu_lmh, cbullied, onlfreq_is, onlpref, famsup_hml, frisup_hml, teachersup, studsup, talkf, talkm, timeexe_r)

#baseline
d_mdl %>%
  group_by(mental_issue) %>%
  tally() %>%
  mutate(prop = n/sum(n))

# data split
set.seed(67)
data_split <- initial_split(d_mdl, prop = 0.8, strata = mental_issue)
train_data <- training(data_split)
test_data  <- testing(data_split)

# 3. Create a Preprocessing Recipe -----------------------------------------
# Define formula (Outcome ~ Predictors) and add preprocessing steps
lr_recipe <- recipe(mental_issue ~ . , 
                    data = train_data) %>%
  step_rm(country_name)
 # step_zv(all_predictors()) %>%          # Remove zero-variance predictors
#  step_normalize(all_numeric_predictors()) # Center and scale numeric variables

# 4. Define the Logistic Regression Model Spec -------------------------------
lr_spec <- logistic_reg() %>%
  set_engine("glm") %>%                  # Use standard generalized linear model
  set_mode("classification")             # Specify binary classification mode

# 5. Create the Workflow ----------------------------------------------------
# Bundle the recipe and model specification together
lr_workflow <- workflow() %>%
  add_recipe(lr_recipe) %>%
  add_model(lr_spec)

# 6. Fit the Model (Train) --------------------------------------------------
# This automatically preps the recipe and trains the model on train data
lr_fit <- fit(lr_workflow, data = train_data)

# 7. Evaluate on Test Data --------------------------------------------------
# Predict classes (labels)
predictions_class <- predict(lr_fit, new_data = test_data)
predictions_class

# Predict probabilities
predictions_prob  <- predict(lr_fit, new_data = test_data, type = "prob")
predictions_prob

# Combine test data with predictions
evaluation_data <- test_data %>%
  bind_cols(predictions_class, predictions_prob)

# Print confusion matrix and accuracy metrics
evaluation_data %>% conf_mat(truth = mental_issue, estimate = .pred_class)
evaluation_data %>% metrics(truth = mental_issue, estimate = .pred_class)

# Extract and print clean model coefficients
lr_fit %>% 
  extract_fit_parsnip() %>% 
  tidy() %>%
  print(n = Inf)

# Print coefficients as Odds Ratios
lr_fit %>% 
  extract_fit_parsnip() %>% 
  tidy(exponentiate = TRUE) %>%
  mutate(is_sig = ifelse(p.value < .05, 1, 0)) %>%
  print(n = Inf)

# Calculate ROC AUC
evaluation_data %>% 
  roc_auc(truth = mental_issue, .pred_0)

# Generate and plot the ROC Curve
evaluation_data %>% 
  roc_curve(truth = mental_issue, .pred_0) %>% 
  autoplot()


```


```{r}
# 1. Isolate just your predictor columns (exclude your target "mental_issue")
# Hint: Use a subset of up to 1,000-2,000 rows for speed
X_explain <- d[1:1000, ] %>% select(-mental_issue)

require(kernelshap)
require(shapviz)

# 2. Compute the SHAP values 
# Type = "prob" ensures R looks at the predicted probabilities (e.g., .pred_Yes)
shap_results <- kernelshap(
  object = lr_fit, 
  X = X_explain, 
  bg_X = d[1:200, ] %>% select(-mental_issue), # Small background baseline
  type = "prob"
)

# 3. Convert to a shapviz object for instant plotting
shp_viz <- shapviz(shap_results)

# Replace ".pred_1" with your positive class prediction column name if different
sv_importance(shp_viz$.pred_1, kind = "beeswarm")

sv_dependence(shp_viz$.pred_1, v = "pmsu_lmh")

for (i in 1:10){
print(sv_waterfall(shp_viz$.pred_1, row_id = i))
}

sv_waterfall(shp_viz$.pred_1, row_id = 5)


```



