---
title: "HBSC"
date: "Updated: `r format(Sys.time(), '%Y-%m-%d')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
urlcolor: blue
---

```{r setup, include=FALSE}

rm(list = ls())

require(tidyverse)
library(knitr)
require(stringr)
library(scales)
library(corrplot)
library(psych)



#prj_fldr <- "C:/MisLocalFiles/Github/HBSC/"
prj_fldr <- "C:/Github/HBSC/"
#setwd(prj_fldr)
#knitr::opts_knit\$set(root.dir = "C:/MisLocalFiles/Github/HBSC")

```

```{r dat_read, echo=FALSE}

dat_fldr <- paste0(prj_fldr,"data/")

hbsc2018 <- read_delim(paste0(dat_fldr, "raw/HBSC2018OAed1.1.csv"),
                       delim = ";",
                       na = c("", " "),
                       show_col_types = FALSE
                       )

```

# Data Overview

It's a huge file:

```{r, echo=FALSE}

dim(hbsc2018)

```

With these columns:

```{r, echo=FALSE}

# get column names
cols <- colnames(hbsc2018)
#cols <- strsplit(cols, ";")[[1]]
#cols <- trimws(cols)

n <- length(cols)
mat <- matrix(cols, ncol = 5, byrow = TRUE)

kable(mat)

```

So after selecting and grouping variables, create dat with :

```{r dat_create, echo=FALSE}

var_groups <- list(
  "Individual Information" = c(
    "seqno_int", "countryno", "weight", "agecat", "sex",
    "MBMI", "IRFAS", "IRRELFAS_LMH", "oweight_who",
    "bodyweight", "bodyheight"
  ),
  "Health Issues and Life Satisfaction" = c(
    "health", "lifesat", "headache", "stomachache", "backache",
    "feellow", "irritable", "nervous", "sleepdificulty", "dizzy"
  ),
  "Problematic Social Media Use (PMSU)" = c(
    "emcsocmed1", "emcsocmed2", "emcsocmed3",
    "emcsocmed4", "emcsocmed5", "emcsocmed6",
    "emcsocmed7", "emcsocmed8", "emcsocmed9"
  ),
  "Bullying" = c(
    "beenbullied", "cbeenbullied"
  ),
  "Communicates Online More Than Real Life" = c(
    "emconlfreq1", "emconlfreq2", "emconlfreq3", "emconlfreq4",
    "emconlpref1", "emconlpref2", "emconlpref3"
  ),
  "Family Support" = c(
    "famhelp", "famsup", "famtalk", "famdec"
  ),
  "Friends Support" = c(
    "friendhelp", "friendcounton", "friendshare", "friendtalk"
  ),
  "School Support" = c(
    "likeschool", "schoolpressure",
    "studtogether", "studhelpful", "studaccept",
    "teacheraccept", "teachercare", "teachertrust"
  ),
  "Communication with Parents" = c(
    "talkfather", "talkmother"
  ),
  "Good/Bad Habits" = c(
    "physact60", "timeexe", "alc30d_2"
  )
)

# create dat
cols_to_keep <- unlist(var_groups, use.names = FALSE)
dat <- hbsc2018 %>%
  select(all_of(cols_to_keep))

# prettt print vars
 for (group_name in names(var_groups)) {
   cat("##", group_name, "\n")
 
   cat(
     paste0(
       "* `", var_groups[[group_name]], "`",
       collapse = "\n"
     ),
     "\n\n"
   )
 }

```

The number of missing values is relevant and need to address later:

```{r nas, echo=FALSE}
na_cnts <- data.frame(
  variable = names(dat),
  na_count = sapply(dat, function(x) sum(is.na(x)))
)

row.names(na_cnts) <- NULL

na_cnts <- na_cnts %>%
  #mutate(na_pct = round(100 * na_count / nrow(dat),0)) %>%
  mutate(na_pct = na_count / nrow(dat)) %>% 
  mutate(na_pct = percent(na_pct, accuracy = 1)) %>%
  arrange(desc(na_count))

print(na_cnts)
```

Now I explore all variables by group and feature engineer as appropriate.

# [3] "Problematic Social Media Use (PMSU)"  
9 binary variables are recoded to 0/1.

```{r pmsu_recode, echo=FALSE}

vars_pmsu <- var_groups[["Problematic Social Media Use (PMSU)"]]
vars_pmsu_r <- paste0(vars_pmsu, "_r")
  
# recode to 0/1 from 1/2
recode_emcsocmed <- function(old_col){
  new_col <- case_when(
    old_col == 1 ~ 0,
    old_col == 2 ~ 1,
    TRUE ~ NA #99?
  )
  return(new_col)
}

dat <- dat %>%
  mutate(emcsocmed1_r = recode_emcsocmed(emcsocmed1),
         emcsocmed2_r = recode_emcsocmed(emcsocmed2),
         emcsocmed3_r = recode_emcsocmed(emcsocmed3),
         emcsocmed4_r = recode_emcsocmed(emcsocmed4),
         emcsocmed5_r = recode_emcsocmed(emcsocmed5),
         emcsocmed6_r = recode_emcsocmed(emcsocmed6),
         emcsocmed7_r = recode_emcsocmed(emcsocmed7),
         emcsocmed8_r = recode_emcsocmed(emcsocmed8),
         emcsocmed9_r = recode_emcsocmed(emcsocmed9)
         ) 

# and look like pprrint:
questions <- data.frame(
  item = c(
    "emcsocmed1","emcsocmed2","emcsocmed3","emcsocmed4",
    "emcsocmed5","emcsocmed6","emcsocmed7","emcsocmed8","emcsocmed9"
  ),
  topic = c(
    "Social media: I can't think of anything else",
    "Social media: Dissatisfied cause want to spend more time",
    "Social media: Felt bad cause want to play",
    "Social media: Failed to spend less time",
    "Social media: Neglected other activities",
    "Social media: Had arguments because of use",
    "Social media: Lied about amount of time spent",
    "Social media: Use it to escape from negative feelings",
    "Social media: Conflict with family because of use"
  )
)

kable(questions) #, caption = "Social Media Use Items")

```

And they are correlated.

```{r corr2,echo=FALSE}

tetra_mat <- tetrachoric(dat[vars_pmsu_r])$rho
tetra_mat <- round(tetra_mat, 2)

corrplot::corrplot(
  tetra_mat,
  method = "color",
  type = "upper",
  tl.col = "black",
  addCoef.col = "black"
)

```

These 9 vars are sumed and then binned to H/M/L where H(6-9), M(2-5) and L(0-1). Resulting in:

```{r pmsu_recode2, echo=FALSE}

#sum and bin
dat <- dat %>%
  # should go from [0,9]
  mutate(pmsu_sum = rowSums(dat[,vars_pmsu_r])) %>%
  # 3 groups
  mutate(
    pmsu_lmh = case_when(
      between(pmsu_sum,6,9) ~ 3, #"High",
      between(pmsu_sum,2,5) ~ 2, #"Med",
      between(pmsu_sum,0,1) ~ 1, #"Low",
      TRUE ~ NA
    )
  )

#cat("## Table for emcsocmed_sum x pmsu_lmh")
kable(table(dat$pmsu_sum,
      dat$pmsu_lmh,
      useNA = "always")
)

# remove original vars
dat <-  dat %>%
  select(-all_of(vars_pmsu))

```

# [6] "Family Support" 
4 questions going from 1 Very strongly disagree to 7 Very strongly agree.

```{r famprnt, echo=FALSE}
family_items <- data.frame(
  item = c(
    "famhelp",
    "famsup",
    "famtalk",
    "famclose"
  ),
  topic = c(
    "Family tries to help",
    "Get emotional help from family",
    "Talk about problems with family",
    "Feel close to family"
  )
)

kable(family_items) #, caption = "Family Support Items")
```

Summing them results in [4,28] and then grouped as H/M/L as L 4-11, M 12-19, H 20-28. Resulting in:

```{r fam_rec, echo=FALSE}

vars_family_support <- c("famhelp", "famsup", "famtalk", "famdec")

dat <- dat %>%
   mutate(famsup_sum = rowSums(dat[,vars_family_support])
          ) %>%
   mutate(famsup_hml = case_when(
            famsup_sum >= 4 & famsup_sum <= 11 ~ 1, #"Low",
            famsup_sum >= 12 & famsup_sum <= 19 ~ 2, #"Med",
            famsup_sum >= 20 & famsup_sum <= 28 ~ 3, #"High",
            TRUE ~ NA))

tbl_famsup <- table(dat$famsup_sum,
                    dat$famsup_hml,
                    useNA = "always"
                    )

kable(tbl_famsup)

```

voy
[1] "Individual Information"                 
[2] "Health Issues and Life Satisfaction"    
[4] "Bullying"                               
[5] "Communicates Online More Than Real Life"
                        
[7] "Friends Support"                        
[8] "School Support"                         
[9] "Communication with Parents"             
[10] "Good/Bad Habits"









