---
title: 'HBSC: Data Overview and Prep'
date: "Updated: `r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    df_print: paged
  word_document: default
  html_notebook: default
  pdf_document: default
urlcolor: blue
---

```{r project_setup}

rm(list = ls())

setwd("C:/MisLocalFiles/Github/HBSC/scripts/data_prep/")
project_folder <- "C:/MisLocalFiles/Github/HBSC/"

require(tidyverse)
require(broom)
require(naniar)
library(purrr)
library(patchwork)
library(knitr)

```

## Dataset Overview

```{r data_load}

hbsc2018 <- read_delim(paste0(project_folder,
                                  "/data/raw/HBSC2018OAed1.1.csv"),
                       delim = ";",
                       na = c("", " "))
# working copy
dat <- hbsc2018

```

File has the following rows and columns:

```{r}

dim(hbsc2018)

```

Pretty much all columns have missing values so need to address later. Though it 
seems much of the variables I will care about don't have much.

```{r}

na_cnts <- data.frame(
  variable = names(dat),
  na_count = sapply(dat, function(x) sum(is.na(x)))
)

row.names(na_cnts) <- NULL

na_cnts <- na_cnts %>%
  mutate(na_pct = round(100 * na_count / nrow(dat),1)) %>%
  arrange(desc(na_count))

print(na_cnts)

```

### seqno_int
This is the unique identifier of each record and I confirmed it's unique.

```{r}

length(unique(dat$seqno_int)) == nrow(dat)

```

### countryno
Number of countries in dataset is:

```{r}

length(unique(dat$countryno))

```
And this is the count of survey responses by country:

```{r country_recode}

recode_map <- c(
  "8000"   = "Albania",
  "31000"  = "Azerbaijan",
  "40000"  = "Austria",
  "51000"  = "Armenia",
  "56001"  = "Belgium (Flemish)",
  "56002"  = "Belgium (French)",
  "100000" = "Bulgaria",
  "124000" = "Canada",
  "191000" = "Croatia",
  "203000" = "Czech Republic",
  "208000" = "Denmark",
  "233000" = "Estonia",
  "246000" = "Finland",
  "250000" = "France",
  "268000" = "Georgia",
  "276000" = "Germany",
  "300000" = "Greece",
  "304000" = "Greenland",
  "348000" = "Hungary",
  "352000" = "Iceland",
  "372000" = "Ireland",
  "376000" = "Israel",
  "380000" = "Italy",
  "398000" = "Kazakhstan",
  "428000" = "Latvia",
  "440000" = "Lithuania",
  "442000" = "Luxembourg",
  "470000" = "Malta",
  "498000" = "Republic of Moldova",
  "528000" = "Netherlands",
  "578000" = "Norway",
  "616000" = "Poland",
  "620000" = "Portugal",
  "642000" = "Romania",
  "643000" = "Russia",
  "688000" = "Serbia",
  "703000" = "Slovakia",
  "705000" = "Slovenia",
  "724000" = "Spain",
  "752000" = "Sweden",
  "756000" = "Switzerland",
  "792000" = "Turkey",
  "804000" = "Ukraine",
  "807000" = "Macedonia",
  "826001" = "England",
  "826002" = "Scotland",
  "826003" = "Wales",
  "826004" = "Northern Ireland",
  "840000" = "USA"
)

dat$country_name <- recode_map[as.character(dat$countryno)]

dat %>%
  group_by(country_name) %>%
  summarise(n = n(),
            .groups = "drop") %>%
  mutate(pct = round(100 * n / sum(n),1)) %>%
  arrange(desc(n)) %>%
  print(n=Inf)

```

## Variables to Predict

### lifesat
Q: Here is a picture of a ladder. The top of the ladder “10” is the best possible life for
you and the bottom “0” is the worst possible life for you. In general, where on the
ladder do you feel you stand at the moment?
A: 0-10 

```{r}

ggplot(dat, aes(y = factor(lifesat))) +
  geom_bar(fill = "steelblue") +
  labs(
    y = "lifesat",
    x = "n"
    ) +
  theme_minimal()

```
lifesat is dichotomized for modeling as lifesat_low:

```{r}

dat <- dat %>%
  mutate(
    lifesat_low = case_when(
      between(lifesat, 0, 5) ~ 1,
      between(lifesat, 6, 10) ~ 0,
      TRUE ~ NA))

table(dat$lifesat,
      dat$lifesat_low,
      useNA = "always"
      )

```
Ultimately the distribution is:

```{r}

dat %>%
  group_by(lifesat_low) %>%
  summarise(n = n()) %>%
  mutate(pct = n / sum(n))

```


### (mental) health complaints 
There are 8 variables for this that reply to a question like: 
In the last 6 months: how often have you had the following….? And the answers goes from
1 About every day to 5 Rarely or never. The 8 variables are: headache, stomachache, backache,
dizzy, feellow, irritable, nervous, sleepdificulty.

Their distributions look like this:

```{r}

vars_health_complaints <- c("headache", "stomachache", "backache", "dizzy",
          "feellow", "irritable", "nervous", "sleepdificulty")

plot_list <- map(vars_health_complaints, function(v) {
  ggplot(dat, aes(y = factor(.data[[v]]))) +
    geom_bar(fill = "steelblue") +
    coord_flip() +
    labs(
      y = v,
      x = "n"
    ) +
    theme_minimal()
})

wrap_plots(plot_list, ncol = 2)

```
These variables are then recoded to a binary where it flags if the complaint
occurs more than once per week (1 or 2).

```{r}

dat <- dat %>%
  mutate(
    headache_frequent = ifelse(headache <= 2, 1, 0),
    stomachache_frequent = ifelse(stomachache <= 2, 1, 0),
    backache_frequent = ifelse(backache <= 2, 1, 0),
    dizzy_frequent = ifelse(dizzy <= 2, 1, 0),
    feellow_frequent = ifelse(feellow <= 2, 1, 0),
    irritable_frequent = ifelse(irritable <= 2, 1, 0),
    nervous_frequent = ifelse(nervous <= 2, 1, 0),
    sleepdificulty_frequent = ifelse(sleepdificulty <= 2, 1, 0)
  )

```

The recoded variables look like this:

```{r}

vars_health_complaints_frequent <- paste0(vars_health_complaints,
                                          "_frequent")

for (i in seq_along(vars_health_complaints)) {
  cat("\n\n###", vars_health_complaints[i], "x", vars_health_complaints_frequent[i], "\n")
  print(table(dat[[vars_health_complaints[i]]], dat[[vars_health_complaints_frequent[i]]], useNA = "always"))
}

```
The frequency of complaints is then added to calculate one sum for physical and 
one for mental. If the sum is 2 or more then it is determined that there is a health issue.
Most papers do it for psycosomatic (all 8 vars) but I will focus on the 4 mental ones.
Still, I include both vars in case.

```{r}

dat <- dat %>%
  # add the frequency of complaints
  mutate(mental_complaints_sum = feellow_frequent +
                           irritable_frequent +
                           nervous_frequent +
                           sleepdificulty_frequent,
         physical_complaints_sum = headache_frequent +
                           stomachache_frequent +
                           backache_frequent + 
                           dizzy_frequent
  ) %>%
  # derive index based on sum
  mutate(mental_complaintsY = ifelse(mental_complaints_sum >=2, 1, 0),
         health_complaintsY = ifelse((mental_complaints_sum + physical_complaints_sum) >=2, 1, 0)
      ) 

cat("\n\n### mental_complaintsY x mental_complaints_sum\n")
table(dat$mental_complaints_sum,
      dat$mental_complaintsY,
      useNA = "always")

cat("\n\n### (health_complaintsY x health_complaints_sum\n")
table(dat$mental_complaints_sum + dat$physical_complaints_sum,
      dat$health_complaintsY,
      useNA = "always")

```

### mental_complaintsY  
Ultimately our variable to predict will be mental_complaintsY; at ~26%.

```{r}

dat %>%
  group_by(mental_complaintsY) %>%
  summarise(n = n()) %>%
  mutate(pct = n / sum(n))

```


## The Predictors

### Problematic social media use (PMSU)
PMSU are the main variables of interest. There are 
9 of them: emcsocmed1 to emcsocmed9. The prequestion for all of them is: 
We are interested in your experiences with social media. The term social
media refers to social network sites (e.g. Facebook, [add other local
examples]) and instant messengers (e.g. [insert local examples],
WhatsApp, Snapchat, Facebook messenger).
During the past year, have you….
Please tick one box for each line.  

The possible answers are 1 (no) and 2 (yes). Also those with 99 mean Missing due to skip pattern. 

The topic of each question is: 

```{r}

questions <- data.frame(
  item = c(
    "emcsocmed1","emcsocmed2","emcsocmed3","emcsocmed4",
    "emcsocmed5","emcsocmed6","emcsocmed7","emcsocmed8","emcsocmed9"
  ),
  topic = c(
    "Social media: I can't think of anything else",
    "Social media: Dissatisfied cause want to spend more time",
    "Social media: Felt bad cause want to play",
    "Social media: Failed to spend less time",
    "Social media: Neglected other activities",
    "Social media: Had arguments because of use",
    "Social media: Lied about amount of time spent",
    "Social media: Use it to escape from negative feelings",
    "Social media: Conflict with family because of use"
  )
)

kable(questions) #, caption = "Social Media Use Items")

```

Since I need to get a sum of them I will recode them from 1/2 to 0/1.These are the 
distributions of the answers. I put the 99 as NA - is that correct?  

The recoding looks like this:

```{r}

# recode emcsocmed to 0/1 from 1/2
recode_emcsocmed <- function(old_col){
  new_col <- case_when(
    old_col == 1 ~ 0,
    old_col == 2 ~ 1,
    TRUE ~ NA #99?
  )
  return(new_col)
}

dat <- dat %>%
  mutate(emcsocmed1_r = recode_emcsocmed(emcsocmed1),
         emcsocmed2_r = recode_emcsocmed(emcsocmed2),
         emcsocmed3_r = recode_emcsocmed(emcsocmed3),
         emcsocmed4_r = recode_emcsocmed(emcsocmed4),
         emcsocmed5_r = recode_emcsocmed(emcsocmed5),
         emcsocmed6_r = recode_emcsocmed(emcsocmed6),
         emcsocmed7_r = recode_emcsocmed(emcsocmed7),
         emcsocmed8_r = recode_emcsocmed(emcsocmed8),
         emcsocmed9_r = recode_emcsocmed(emcsocmed9)
         ) 

vars <- paste0("emcsocmed", 1:9)

for (v in vars) {
  cat("## Table for", v)
  print(table(dat[[v]], dat[[paste0(v, "_r")]], useNA = "always"))
  cat("\n\n")
}

```

And this gives an idea of how correlated they are with each other. 

Is this right? Too correlated? Or it's done different? It has to be done another way?

```{r}

vars <- paste0("emcsocmed", 1:9)

r <- cor(dat[,vars],
    use = "complete.obs")

round(r,4)

```

PMSU variables are then combined so output is represented in one variable, instead of 9.
I will do 2 versions:
a) Yes/No where Y(6-9) and N(0-5)
b) H/M/L where H(6-9), M(2-5) and L(0-1)

The distribution of PMSU categorized is:

```{r}

vars_emcsocmed <- paste0("emcsocmed", 1:9)
vars_emcsocmed_r <- paste0("emcsocmed", 1:9, "_r")

dat <- dat %>%
  mutate(emcsocmed_sum = rowSums(dat[,vars_emcsocmed_r])) %>%
  mutate(
    pmsu_yn = case_when(
      between(emcsocmed_sum,6,9) ~ 1,
      between(emcsocmed_sum,0,5) ~ 0,
      TRUE ~ NA
    ),
    pmsu_lmh = case_when(
      between(emcsocmed_sum,6,9) ~ "High",
      between(emcsocmed_sum,2,5) ~ "Med",
      between(emcsocmed_sum,0,1) ~ "Low",
      TRUE ~ NA
    )
  )

cat("## Table for emcsocmed_sum x pmsu_yn")
table(dat$emcsocmed_sum,
      dat$pmsu_yn,
      useNA = "always")

cat("## Table for emcsocmed_sum x pmsu_lmh")
table(dat$emcsocmed_sum,
      dat$pmsu_lmh,
      useNA = "always")

cat("## Table for pmsu_yn x pmsu_lmh")
table(dat$pmsu_lmh, dat$pmsu_yn, useNA = "always")

```

### Family Support
There are 4 of them. The question starts with: 
We are interested in how you feel about the following statements.
Please show how much you agree or disagree with each one.
Please tick one box for each line.

The options to answer go from 1 Very strongly disagree to 7 Very strongly agree.
The 4 questions are: 

```{r}

family_items <- data.frame(
  item = c(
    "famhelp",
    "famsup",
    "famtalk",
    "famclose"
  ),
  topic = c(
    "Family tries to help",
    "Get emotional help from family",
    "Talk about problems with family",
    "Feel close to family"
  )
)

kable(family_items) #, caption = "Family Support Items")

```

These variables are then combined to create a categorical predictor. For now,
I will keep two options like I did for PMSU but will select only one when modeling.
These are:  
a) H/L based on avg of 4 vars. L if [1,5.5) and H if [5.5, 7]  
b) H/M/L based on sum of 4 vars. L if 4-11, M if 12-19, H if 20-28.

```{r}

vars_family_support <- c("famhelp", "famsup", "famtalk", "famdec")

dat <- dat %>%
   mutate(family_support_avg = rowMeans(dat[,vars_family_support]),
          family_support_sum = rowSums(dat[,vars_family_support])
          ) %>%
   mutate(family_support_high = case_when(
            family_support_avg >= 1 & family_support_avg < 5.5 ~ 0,
            family_support_avg >= 5.5 & family_support_avg <= 7 ~ 1,
            TRUE ~ NA)) %>%
   mutate(family_support_hml = case_when(
            family_support_sum >= 4 & family_support_sum <= 11 ~ "Low",
            family_support_sum >= 12 & family_support_sum <= 19 ~ "Med",
            family_support_sum >= 20 & family_support_sum <= 28 ~ "High",
            TRUE ~ NA))

```

Their counts after recoding are:  

```{r}
# #table(dat$family_support_high, dat$family_support_hml, useNA = "always")
# cat("## Table for family_support_high")
# table(dat$family_support_high, useNA = "always")

#cat("## Table for family_support_hml")
#table(dat$family_support_hml, useNA = "always")

t <- data.frame(
          table(dat$family_support_high, useNA = "always")
          ) %>%
  mutate(pct = round(Freq / sum(Freq),3))

kable(t, caption = "family_support_high")

t <- data.frame(
          table(dat$family_support_hml, useNA = "always")
          ) %>%
  mutate(pct = round(Freq / sum(Freq),3))

kable(t, caption = "family_support_hml")

```

### Friends (Peer) Support  
There are 4 friends support questions. The questions start with: 
We are interested in how you feel about the following statements.
Please show how much you agree or disagree with each one.
Please tick one box for each line.  

The answers range from 1 Very strongly disagree to 7 Very strongly agree.  

```{r}
friend_items <- data.frame(
  item = c(
    "friendhelp",
    "friendcounton",
    "friendshare",
    "friendtalk"
  ),
  topic = c(
    "My friends really try to help me",
    "I can count on my friends when things go wrong",
    "I have friends with whom I can share my joys and sorrows",
    "I can talk about my problems with my friends"
  )
)

kable(friend_items)

```

  
These variables are then combined to create a categorical predictor. For now,
I will keep two options like I did for family but will select only one when modeling.
These are:  
a) H/L based on avg of 4 vars. L if [1,5.5) and H if [5.5, 7]  
b) H/M/L based on sum of 4 vars. L if 4-11, M if 12-19, H if 20-28.


```{r}

vars_friends_support <- c("friendhelp", "friendcounton", "friendshare", "friendtalk")
# recode friends/peer support as yn and hml
dat <- dat %>%
  mutate(friends_support_avg = rowMeans(dat[,vars_friends_support]),
         friends_support_sum = rowSums(dat[,vars_friends_support])
  ) %>%
  mutate(friends_support_high = case_when(
    friends_support_avg >= 1 & friends_support_avg < 5.5 ~ 0,
    friends_support_avg >= 5.5 & friends_support_avg <= 7 ~ 1,
    TRUE ~ NA)) %>%
  mutate(friends_support_hml = case_when(
    friends_support_sum >= 4 & friends_support_sum <= 11 ~ "Low",
    friends_support_sum >= 12 & friends_support_sum <= 19 ~ "Med",
    friends_support_sum >= 20 & friends_support_sum <= 28 ~ "High",
    TRUE ~ NA))
```

Their counts after recoding are:  

```{r}

t <- data.frame(
          table(dat$friends_support_high, useNA = "always")
          ) %>%
  mutate(pct = round(Freq / sum(Freq),3))

kable(t, caption = "friends_support_high")

t <- data.frame(
          table(dat$friends_support_hml, useNA = "always")
          ) %>%
  mutate(pct = round(Freq / sum(Freq),3))

kable(t, caption = "friends_support_hml")

# table(dat$friends_support_high, dat$friends_support_hml, useNA = "always")
# table(dat$friends_support_high, useNA = "always")
# table(dat$friends_support_hml, useNA = "always")

```


### School support  
This metric is a combination of multiple questions from 2 categories: teacher
acceptance and classmates acceptance. The answers to these
questions go from 1 (strong agree) to 5 (strong disagree).  

The questions are:  

```{r}

school_items <- data.frame(
  category = c("teacher", "teacher", "teacher",
               "classmates", "classmates", "classmates"),
  item = c("teacheraccept",
           "teachercare",
           "teachertrust",
           "studtogether",
           "studhelpful",
           "studaccept"),
  topic = c("Teacher accepts me",
            "Teacher cares about me",
            "Feel trust in teacher",
            "Students enjoy being together",
            "Students kind and helpful",
            "Students accept me")
)

kable(school_items)

```

The scales of these variables is reversed so 1 is negative and 5 is the most positive. 
Then an arithmetic average is calculated for teacher and students. 
The based on this average a high/low support is calculated for each with 4 being the threshold.

```{r}

# recodes from 1-5 to 0-4 (06-italy)
recode_school_vars <- function(old_values){
  
  new_values <- case_when(
    old_values == 5 ~ 1,
    old_values == 4 ~ 2,
    old_values == 3 ~ 3,
    old_values == 2 ~ 4,
    old_values == 1 ~ 5,
    TRUE ~ NA
  )
  
  return(new_values)
  
}

dat <- dat %>%
  # recode they can be added
  mutate(teacheraccept_r = recode_school_vars(teacheraccept),
         teachercare_r = recode_school_vars(teachercare),
         teachertrust_r = recode_school_vars(teachertrust),
         studtogether_r = recode_school_vars(studtogether),
         studhelpful_r = recode_school_vars(studhelpful),
         studaccept_r = recode_school_vars(studaccept)
         ) %>%
  # derive avg by teacher/student
  mutate(teacher_support_avg = (teacheraccept_r 
                               + teachercare_r
                               + teachertrust_r) / 3,
         student_support_avg = (studtogether_r
                                + studhelpful_r
                                + studaccept_r) / 3,
         # just for fun combine and try
         school_support_avg = (teacheraccept_r 
                                + teachercare_r
                                + teachertrust_r
                                + studtogether_r
                                + studhelpful_r
                                + studaccept_r) / 6,
         ) %>%
  # derive support yn based on hbsc
  mutate(
    teacher_support_high = case_when(
      teacher_support_avg >= 4 ~ 1,
      teacher_support_avg < 4 ~ 0,
      TRUE ~ NA),
      student_support_high = case_when(
        student_support_avg >= 4 ~ 1,
        student_support_avg < 4 ~ 0,
        TRUE ~ NA),
      school_support_high = case_when(
        school_support_avg >= 4 ~ 1,
        school_support_avg < 4 ~ 0,
      TRUE ~ NA)
  )

```

The recoding of each var is:

```{r}

vars_school <- c("teacheraccept",
           "teachercare",
           "teachertrust",
           "studtogether",
           "studhelpful",
           "studaccept")

vars_school_r <- paste0(vars_school,
                        "_r")

for (i in seq_along(vars_school)) {
  cat("\n\n###", vars_school[i], "x", vars_school_r[i], "\n")
  print(table(dat[[vars_school[i]]], dat[[vars_school_r[i]]], useNA = "always"))
}

```

And the actual recoded predictors look like:

```{r}

t <- data.frame(table(dat$teacher_support_high, useNA = "always")) %>%
  mutate(pct = round(Freq / sum(Freq),3))
kable(t, caption = "teacher_support_high")

t <- data.frame(table(dat$student_support_high, useNA = "always")) %>%
  mutate(pct = round(Freq / sum(Freq),3))
kable(t, caption = "student_support_high")

t <- data.frame(table(dat$school_support_high, useNA = "always")) %>%
  mutate(pct = round(Freq / sum(Freq),3))
kable(t, caption = "school_support_high")

```


### IRRELFAS_LMH: Relative family affluence categorical  
Derived from IRFAS. So check list of fas-variables in the dataset. 
It is a cut based on percentiles so I recode to friendly.
1 Lowest 20 pct  
2 Medium 60 pct  
3 Highest 20 pct  

```{r}

dat <- dat %>%
  mutate(IRRELFAS_LMH_r = 
           case_when(IRRELFAS_LMH == 1 ~ "Low20",
                     IRRELFAS_LMH == 2 ~ "Med60",
                     IRRELFAS_LMH == 3 ~ "High20",
                     TRUE ~ NA)
  )

t <- table(dat$IRRELFAS_LMH,
           dat$IRRELFAS_LMH_r,
           useNA = "always")

kable(t,
       caption = "IRRELFAS_LMH x IRRELFAS_LMH_r")

```


### talkfather and talkmother
Two questions measure parent communication. The pre-question is:
How easy is it for you to talk to the following persons about things that
really bother you?
Please tick one box for each line  

Answers go from 1 Very easy to 5 Don't have or see.  

These are dichotomized for each parent as Yes if [1,2] and No if [3-5].  

```{r}

dat <- dat %>%
  mutate(talkfatherYes = 
           case_when(talkfather %in% c(1,2) ~ 1,
                     talkfather %in% c(3,4,5) ~ 0,
                     TRUE ~ NA),
         talkmotherYes = 
           case_when(talkmother %in% c(1,2) ~ 1,
                     talkmother %in% c(3,4,5) ~ 0,
                     TRUE ~ NA)
         )

cat("## Table for talkfatherYes x talkfather")
table(dat$talkfather, dat$talkfatherYes, useNA = "always")

cat("## Table for talkmotherYes x talkmother")
table(dat$talkmother, dat$talkmotherYes, useNA = "always")

```
I was going to combine to a "number of parents one can talk" but I'll leave for now
and will try as interactions later.

### bullying victim  
There are 2 variables that measure if kid has been bullied. a) beenbullied: Been bullied past months
and b) cbeenbullied: Been cyber bullied. Answers range from 1 Haven't to 5 Several times a week.  

These are dichotomized as Yes if [2-5] and No if [1].

```{r}

dat <- dat %>%
  mutate(beenbulliedYes = 
           case_when(beenbullied == 1 ~ 0,
                     between(beenbullied, 2, 5) ~ 1,
                     TRUE ~ NA),
         cbeenbulliedYes = 
           case_when(cbeenbullied == 1 ~ 0,
                     between(cbeenbullied, 2, 5) ~ 1,
                     TRUE ~ NA)
           )

cat("## Table for beenbullied x beenbulliedYes")
table(dat$beenbullied, dat$beenbulliedYes, useNA = "always")

cat("## Table for cbeenbullied x cbeenbulliedYes")
table(dat$cbeenbullied, dat$cbeenbulliedYes, useNA = "always")

```

### age and sex(gender)  
There are 3 age groups and 2 genders. I recode to friendly/actual value and they look
like this.  

```{r}

dat <- dat %>%
  mutate(
    age_r = case_match(agecat,
                             1 ~ 11,
                             2 ~ 13,
                             3 ~ 15,
                             .default = NA
                             ),
    sex_r = case_match(sex,
                             1 ~ "Boy",
                             2 ~ "Girl",
                             .default = NA)
  )

cat("## Table for agecat x age_r")
table(dat$agecat, dat$age_r, useNA = "always")

cat("## Table for sex x sex_r")
table(dat$sex, dat$sex_r, useNA = "always")

cat("## Table for age_r x sex_r")
table(dat$age_r, dat$sex_r, useNA = "always")

```

### Continuos Online Communications
This category measures how how often young person communicate online with 
4 types of people in their life. Preamble of question is: How often do you have ONLINE 
contact with the following people?
Answers range from 1 (NA or don't know) to 6 (almost all of the time). 

| Variable      | Description                         |
|---------------|-------------------------------------|
| emconlfreq1   | Online contact with close friends    |
| emconlfreq2   | Online contact with larger friend group |
| emconlfreq3   | Online contact with online friends   |
| emconlfreq4   | Online contact with other            |

I combine this into one dichotomized and adopt the HBSC definition of yes if 
any of the variables is 6. 

```{r}

dat <- dat %>%
  mutate(emconlfreqYes = case_when(
    (emconlfreq1 == 6 | emconlfreq2 == 6 | emconlfreq3 == 6 | emconlfreq4 == 6) ~ 1,
    (!is.na(emconlfreq1) | !is.na(emconlfreq2) | !is.na(emconlfreq3) | !is.na(emconlfreq4)) ~ 0,
    TRUE ~ NA
  )) 
#still need to do something with 99

table(dat$emconlfreqYes, useNA = "always")

```

### Prefer Online life
I haven't found a paper that include this but I want to. It asks questions about
preference of sharing important stuff online vs real-life. The three variables are:

| Variable      | Description                     |
|---------------|---------------------------------|
| emconlpref1   | Secrets, more easily online     |
| emconlpref2   | Feelings, more easily online    |
| emconlpref3   | Concerns, more easily online    |

The answers range from 1 (Strong disagree) to 5 (Strong agree).
The counts are:

```{r}

online_life <- vars <- c("emconlpref1", "emconlpref2", "emconlpref3")

for (i in seq_along(online_life)){
  #print(online_life[i])
  #print(table(dat[online_life[1]], useNA = "ifany"))
  
  print(
    dat %>% 
      group_by(.data[[ online_life[i] ]]) %>% 
      tally() %>%
      mutate(pct = round(n/sum(n),3))
  )
}

```

For this I copy the school approach which is avg and dichotomize with 4 as threshold:  

```{r}

dat <- dat %>%
  mutate(lives_online = (emconlpref1 +  emconlpref2 +  emconlpref3)/3) %>%
  mutate(lives_onlineYN = ifelse(lives_online >= 4, 1, 0)
         )

table(dat$lives_online %/% 1,
      dat$lives_onlineYN,
      useNA = "always")

```

# timeexe: Vigorous physical activity frequency
Question is Outside school hours: how often do you usually exercise in your free time so much
that you get out of breath or sweat? Answers go from 1 (every day) to 7 never. 

HBSC shows proportions of those that do three or more times per week. But that is 2022 data,
in our case we do 2-3 times per week or more.

```{r}

dat <- dat %>%
  mutate(timeexe_r = case_when(
    between(timeexe, 1, 3) ~ 1,
    between(timeexe, 4, 7) ~ 0,
    TRUE ~ NA
  ))

print(
table(dat$timeexe,
      dat$timeexe_r,
      useNA = "ifany")
)

```

# alc30d_2: Alcohol use last 30 days
Question is  On how many days (if any) have you drunk alcohol?
Answers are binned and go from 1 (never) to 7 (30 days or more).
HBSC dichotomizes as Never vs Yes so I will do the same.

```{r}

dat <- dat %>%
  mutate(alc30d_YN = case_when(
    alc30d_2 == 1 ~ 0,
    alc30d_2 != 1 ~ 1,
    TRUE ~ NA
  ))

print(
  table(dat$alc30d_2,
        dat$alc30d_YN,
        useNA = "ifany")
)

```


## Dataset for Modeling  
Only the relevant and prepared variables are exported to a csv so I can start the 
modeling process with those. File goes to "data/processed/" and includes date.

```{r}

dat_out <- dat %>%
  select(seqno_int,
         country_name,
         lifesat_low, 
         mental_complaintsY, 
         health_complaintsY, 
         pmsu_lmh, 
         pmsu_yn, 
         family_support_high, 
         family_support_hml, 
         friends_support_high, 
         friends_support_hml, 
         teacher_support_high,
         student_support_high,
         school_support_high,
         IRRELFAS_LMH_r, 
         talkfatherYes, 
         talkmotherYes, 
         beenbulliedYes, 
         cbeenbulliedYes, 
         age_r, 
         sex_r,
         emconlfreqYes,
         lives_onlineYN,
         timeexe_r,
         alc30d_YN
         )

summary(dat_out)

dat_out_filename <- paste0(project_folder,
                           "data/processed/",
                           "dat_", Sys.Date(), ".csv")

#write_csv(dat_out,
 #         dat_out_filename)

# output el full mejor
write_csv(dat,
          dat_out_filename)

```




